---
params:
  data_source: "local"
format:
  dashboard:
    orientation: rows
    scrolling: true
    theme:
      - cosmo
      - custom.scss
execute:
  echo: false
  message: false
  warning: false
---

```{r}
#| label: setup
#| include: false

library(sf)
library(leaflet)
library(dplyr)
library(DT)
library(gt)

# ── Package loading ────────────────────────────────────────────────────────────
# Try pkgload first (dev mode), then library (installed), then graceful fail.
tryCatch(
  pkgload::load_all(quiet = TRUE),
  error = function(e) {
    tryCatch(library(acd), error = function(e2) NULL)
  }
)

# Snapshot date constant (defined in R/snapshot.R; fallback if not loaded)
if (!exists("SNAPSHOT_DATE")) SNAPSHOT_DATE <- "2026-05-01"

# ── Data source selection (Stage 2: "remote" uses GitHub Pages Parquet) ────────
# Default: "local" — reads from inst/extdata (fast, offline, always available).
# Set to "remote" to test the duckplyr path against GitHub Pages Parquet.
# Can also be overridden via params: in YAML front-matter for testing.
data_source <- if (!is.null(params$data_source)) params$data_source else "local"

# ── Parquet cache helper ───────────────────────────────────────────────────────
# Resolves the inst/extdata path for both installed and dev-tree usage,
# then reads via arrow::read_parquet().
.cached_parquet <- function(filename) {
  installed <- system.file("extdata", filename, package = "acd")
  if (nzchar(installed) && file.exists(installed)) {
    return(arrow::read_parquet(installed))
  }
  dev_path <- tryCatch(
    file.path(find.package("acd"), "inst", "extdata", filename),
    error = function(e) ""
  )
  if (nzchar(dev_path) && file.exists(dev_path)) {
    return(arrow::read_parquet(dev_path))
  }
  vignette_dir <- tryCatch(
    dirname(normalizePath(knitr::current_input())),
    error = function(e) ""
  )
  if (nzchar(vignette_dir)) {
    src_path <- normalizePath(
      file.path(vignette_dir, "..", "..", "inst", "extdata", filename),
      mustWork = FALSE
    )
    if (file.exists(src_path)) return(arrow::read_parquet(src_path))
  }
  NULL
}

# ── Placeholder data (used when real parquet is absent) ────────────────────────
# Creates a synthetic sf with all 6 link_status values so the dashboard
# renders completely before real WFS data arrives.
.placeholder_joined <- function() {
  set.seed(42L)
  n <- 18L

  # Vienna centre coordinates + small offsets (degrees)
  lat_c <- 48.2082
  lng_c <- 16.3725

  lats <- lat_c + runif(n, -0.03, 0.03)
  lngs <- lng_c + runif(n, -0.04, 0.04)

  # Create point geometries in WGS84 (for placeholder — display only)
  geom <- sf::st_sfc(
    lapply(seq_len(n), function(i) sf::st_point(c(lngs[i], lats[i]))),
    crs = 4326L
  )

  link_statuses <- factor(
    c(
      rep("in_zone",            7L),
      rep("outside_any_zone",   4L),
      rep("straddles_boundary", 3L),
      rep("multiple_zones",     2L),
      rep("orphan_zone",        1L),
      rep("invalid_geometry",   1L)
    ),
    levels = c(
      "in_zone", "outside_any_zone", "straddles_boundary",
      "multiple_zones", "orphan_zone", "invalid_geometry"
    )
  )

  zone_ids <- c(
    rep("ZONE-A1", 4L), rep("ZONE-B2", 3L), rep(NA_character_, 4L),
    rep("ZONE-A1", 3L), rep(c("ZONE-A1", "ZONE-B2"), 1L),
    "ZONE-C3", NA_character_
  )

  zone_types <- dplyr::case_when(
    link_statuses == "outside_any_zone" ~ NA_character_,
    zone_ids == "ZONE-A1" ~ "district_heating",
    zone_ids == "ZONE-B2" ~ "heat_pump",
    zone_ids == "ZONE-C3" ~ "fossil_fuel_ban",
    TRUE ~ NA_character_
  )

  legal_statuses <- dplyr::case_when(
    zone_types == "district_heating" ~ "district_heating_mandatory",
    zone_types == "heat_pump"        ~ "heat_pump_required",
    zone_types == "fossil_fuel_ban"  ~ "fossil_fuel_banned",
    TRUE ~ NA_character_
  )

  district_labels <- c(
    rep("01", 4L), rep("02", 4L), rep("10", 4L),
    rep("13", 3L), rep("22", 3L)
  )

  sf::st_sf(
    building_id         = paste0("B-SYNTH-", sprintf("%03d", seq_len(n))),
    district            = district_labels,
    year_built          = as.integer(sample(1880:2020, n, replace = TRUE)),
    height_m            = round(runif(n, 8, 40), 1),
    address             = paste0("Musterstrasse ", seq_len(n), ", Wien"),
    zone_id             = zone_ids,
    zone_type           = zone_types,
    legal_status        = legal_statuses,
    link_status         = link_statuses,
    n_overlapping_zones = dplyr::case_when(
      link_statuses == "outside_any_zone"   ~ 0L,
      link_statuses == "multiple_zones"     ~ 2L,
      link_statuses == "straddles_boundary" ~ 1L,
      TRUE ~ 1L
    ),
    overlap_fraction    = dplyr::case_when(
      link_statuses == "in_zone"            ~ 0.98,
      link_statuses == "outside_any_zone"   ~ 0.00,
      link_statuses == "straddles_boundary" ~ round(runif(
        sum(link_statuses == "straddles_boundary"), 0.10, 0.90), 2),
      link_statuses == "multiple_zones"     ~ 0.70,
      TRUE ~ 0.00
    ),
    geometry = geom
  )
}

# ── Placeholder zones (3 fake zones as polygons) ───────────────────────────────
.placeholder_zones <- function() {
  lat_c <- 48.2082
  lng_c <- 16.3725

  make_box <- function(lat0, lng0, dlat = 0.012, dlng = 0.016) {
    sf::st_polygon(list(matrix(
      c(lng0,        lat0,
        lng0 + dlng, lat0,
        lng0 + dlng, lat0 + dlat,
        lng0,        lat0 + dlat,
        lng0,        lat0),
      ncol = 2, byrow = TRUE
    )))
  }

  geom <- sf::st_sfc(
    make_box(lat_c - 0.01,  lng_c - 0.02),
    make_box(lat_c + 0.005, lng_c + 0.005),
    make_box(lat_c - 0.025, lng_c + 0.01),
    crs = 4326L
  )

  sf::st_sf(
    zone_id      = c("ZONE-A1", "ZONE-B2", "ZONE-C3"),
    zone_type    = c("district_heating", "heat_pump", "fossil_fuel_ban"),
    legal_status = c(
      "district_heating_mandatory",
      "heat_pump_required",
      "fossil_fuel_banned"
    ),
    area_ha      = c(12.4, 8.7, 5.1),
    geometry     = geom
  )
}

# ── link_status summary helper (inline fallback if acd package not loaded) ─────
.link_status_summary_local <- function(joined) {
  joined |>
    sf::st_drop_geometry() |>
    dplyr::count(link_status, name = "n_buildings") |>
    dplyr::mutate(
      pct = round(100 * n_buildings / sum(n_buildings), 1)
    ) |>
    dplyr::arrange(link_status)
}

# ── Load or synthesise data ────────────────────────────────────────────────────
# Stage 1: try load_acd_data() (package function); fall back to .cached_parquet;
# fall back to placeholder when neither source is available.
joined_raw <- tryCatch(
  {
    df <- acd::load_acd_data("joined", source = data_source)
    df
  },
  error = function(e) {
    message("load_acd_data() failed (", conditionMessage(e), "); trying .cached_parquet fallback")
    .cached_parquet(paste0("joined_", SNAPSHOT_DATE, ".parquet"))
  }
)

using_placeholder <- is.null(joined_raw)

if (using_placeholder) {
  joined <- .placeholder_joined()
} else {
  # Convert to sf if parquet stripped the geometry column
  if (!inherits(joined_raw, "sf")) {
    joined <- sf::st_as_sf(joined_raw, wkt = "geometry_wkt", crs = 31287L)
  } else {
    joined <- joined_raw
  }
  # Transform to WGS84 for leaflet
  joined <- sf::st_transform(joined, crs = 4326L)
  # Add compatibility aliases for columns renamed in the real schema
  if (!"zone_type" %in% names(joined) && "acd_native" %in% names(joined)) {
    joined$zone_type <- as.character(joined$acd_native)
  }
  if (!"legal_status" %in% names(joined) && "acd_agreement" %in% names(joined)) {
    joined$legal_status <- as.character(joined$acd_agreement)
  }
  if (!"height_m" %in% names(joined) && "n_stories" %in% names(joined)) {
    joined$height_m <- as.numeric(joined$n_stories)
  }
}

# Zones: derive from joined or load separately
zones_raw <- .cached_parquet(paste0("zones_", SNAPSHOT_DATE, ".parquet"))
if (is.null(zones_raw)) {
  zones <- .placeholder_zones()
} else {
  if (!inherits(zones_raw, "sf")) {
    zones <- sf::st_as_sf(zones_raw, wkt = "geometry_wkt", crs = 31287L)
  } else {
    zones <- zones_raw
  }
  # Add canonical zone_id from ERPLABEL if not already present
  if (!"zone_id" %in% names(zones) && "ERPLABEL" %in% names(zones)) {
    zones$zone_id <- zones$ERPLABEL
  }
  # Add zone_type alias from acd_native if present on zones (else use placeholder)
  if (!"zone_type" %in% names(zones)) {
    zones$zone_type <- NA_character_
  }
  # Add legal_status alias
  if (!"legal_status" %in% names(zones)) {
    zones$legal_status <- NA_character_
  }
  zones <- sf::st_transform(zones, crs = 4326L)
}

# Buildings: subset of joined that are actual building points/polygons
buildings <- joined

# ── Pre-compute summary objects ────────────────────────────────────────────────
# link_status summary — use package function if available, else local fallback
link_summary <- tryCatch(
  acd:::link_status_summary(joined),
  error = function(e) .link_status_summary_local(joined)
)

# Counts by link_status for inline prose
n_total      <- nrow(joined)
n_zones      <- nrow(zones)

ls_counts <- link_summary |>
  dplyr::select(link_status, n_buildings) |>
  tibble::deframe()

n_in_zone     <- ls_counts[["in_zone"]]            %||% 0L
n_outside     <- ls_counts[["outside_any_zone"]]   %||% 0L
n_straddle    <- ls_counts[["straddles_boundary"]] %||% 0L
n_multi       <- ls_counts[["multiple_zones"]]     %||% 0L
n_invalid     <- ls_counts[["invalid_geometry"]]   %||% 0L

pct_outside   <- round(100 * n_outside  / n_total, 1)
pct_straddle  <- round(100 * n_straddle / n_total, 1)
pct_multi     <- round(100 * n_multi    / n_total, 1)
pct_in_zone   <- round(100 * n_in_zone  / n_total, 1)

# Orphan zones (zones-side)
orphan_zones <- tryCatch(
  {
    bldgs_31287 <- if (sf::st_crs(buildings)$epsg == 4326L) {
      sf::st_transform(buildings, crs = 31287L)
    } else {
      buildings
    }
    zones_31287 <- if (sf::st_crs(zones)$epsg == 4326L) {
      sf::st_transform(zones, crs = 31287L)
    } else {
      zones
    }
    acd:::find_orphan_zones(zones_31287, bldgs_31287) |>
      sf::st_transform(crs = 4326L)
  },
  error = function(e) {
    # Local fallback: zones not covered by any in_zone or straddles building
    covered_zones <- joined |>
      sf::st_drop_geometry() |>
      dplyr::filter(!is.na(zone_id)) |>
      dplyr::pull(zone_id) |>
      unique()
    zones[!zones$zone_id %in% covered_zones, ]
  }
)
n_orphan <- nrow(orphan_zones)

# ── ACD agreement pre-computation ─────────────────────────────────────────────
# Confusion matrix: rows = spatial says in zone, cols = city says in zone.
# Uses acd_agreement_summary() if available, else local fallback.
.acd_agreement_summary_local <- function(df) {
  df |>
    sf::st_drop_geometry() |>
    dplyr::count(acd_agreement, name = "n_buildings") |>
    dplyr::mutate(pct = round(100 * n_buildings / sum(n_buildings), 1))
}

acd_agree_sum <- tryCatch(
  acd:::acd_agreement_summary(joined),
  error = function(e) .acd_agreement_summary_local(joined)
)

# Disaggregated counts for inline prose (dynamic, no hardcoding)
.ag <- function(level) {
  v <- acd_agree_sum$n_buildings[acd_agree_sum$acd_agreement == level]
  if (length(v) == 0L) 0L else as.integer(v)
}
n_agree_in_zone  <- .ag("agree_in_zone")
n_agree_no_zone  <- .ag("agree_no_zone")
n_spatial_only   <- .ag("spatial_only")
n_acd_only       <- .ag("acd_only")
n_agree_total    <- n_agree_in_zone + n_agree_no_zone
n_disagree_total <- n_spatial_only + n_acd_only

# Confusion matrix as a 2×2 data frame
.cm_cell <- function(level) {
  v <- acd_agree_sum$n_buildings[acd_agree_sum$acd_agreement == level]
  pv <- acd_agree_sum$pct[acd_agree_sum$acd_agreement == level]
  n  <- if (length(v)  == 0L) 0L  else as.integer(v)
  p  <- if (length(pv) == 0L) 0.0 else as.numeric(pv)
  paste0(n, " (", p, "%)")
}
confusion_matrix_df <- data.frame(
  Spatial_join = c("Spatial: IN zone", "Spatial: NOT in zone"),
  City_ACD_yes = c(.cm_cell("agree_in_zone"), .cm_cell("acd_only")),
  City_ACD_no  = c(.cm_cell("spatial_only"),  .cm_cell("agree_no_zone")),
  check.names  = FALSE,
  stringsAsFactors = FALSE
)
colnames(confusion_matrix_df) <- c(
  "Spatial join result",
  "City ACD: yes (non-empty)",
  "City ACD: no (empty/NA)"
)

# Disagreement subset for leaflet
acd_disagree_sf <- tryCatch(
  acd:::case_acd_disagreement(joined),
  error = function(e) {
    # Fallback: filter directly if the package function is unavailable
    if ("acd_agreement" %in% names(joined)) {
      joined[joined$acd_agreement %in% c("spatial_only", "acd_only"), ]
    } else {
      joined[integer(0), ]
    }
  }
)

# Colors for the disagreement leaflet
disagree_colors <- c(
  spatial_only = "#e7298a",   # pink
  acd_only     = "#66a61e"    # olive
)
pal_disagree <- leaflet::colorFactor(
  palette  = unname(disagree_colors),
  levels   = names(disagree_colors),
  na.color = "#999999"
)

# ── Per-case subsets for page 4 ───────────────────────────────────────────────
case_in_zone <- function(df, n = 50L) {
  sub <- df[df$link_status == "in_zone", ]
  if (nrow(sub) > n) sub <- sub[sample(nrow(sub), n), ]
  sub
}

case_outside <- function(df, n = 100L) {
  sub <- df[df$link_status == "outside_any_zone", ]
  if (nrow(sub) > n) sub <- sub[sample(nrow(sub), n), ]
  sub
}

case_straddle <- function(df, n = 50L) {
  sub <- df[df$link_status == "straddles_boundary", ]
  if (nrow(sub) > n) sub <- sub[sample(nrow(sub), n), ]
  sub
}

case_multi <- function(df, n = 50L) {
  sub <- df[df$link_status == "multiple_zones", ]
  if (nrow(sub) > n) sub <- sub[sample(nrow(sub), n), ]
  sub
}

case_invalid <- function(df) {
  df[df$link_status == "invalid_geometry", ]
}

# ── Color palette for link_status ─────────────────────────────────────────────
link_status_colors <- c(
  in_zone            = "#2ca02c",  # green
  outside_any_zone   = "#7f7f7f",  # gray
  straddles_boundary = "#ff7f0e",  # orange
  multiple_zones     = "#9467bd",  # purple
  orphan_zone        = "#1f77b4",  # blue
  invalid_geometry   = "#d62728"   # red
)

link_status_labels <- c(
  in_zone            = "Clean join (in zone)",
  outside_any_zone   = "Outside any zone",
  straddles_boundary = "Straddles boundary",
  multiple_zones     = "Multiple zones",
  orphan_zone        = "Orphan zone",
  invalid_geometry   = "Invalid geometry"
)

# Leaflet color function for link_status
pal_link <- leaflet::colorFactor(
  palette = unname(link_status_colors),
  levels  = names(link_status_colors),
  na.color = "#999999"
)

# Color function for zone_type
zone_type_colors <- c(
  district_heating = "#e41a1c",
  heat_pump        = "#377eb8",
  fossil_fuel_ban  = "#984ea3"
)
pal_zone_type <- leaflet::colorFactor(
  palette  = unname(zone_type_colors),
  levels   = names(zone_type_colors),
  na.color = "#aaaaaa"
)

# Continuous palette for year_built
year_range <- range(buildings$year_built, na.rm = TRUE)
pal_year   <- leaflet::colorNumeric(
  palette = "YlOrRd",
  domain  = year_range
)

# ── Vienna map centre ─────────────────────────────────────────────────────────
vienna_center <- c(lng = 16.3725, lat = 48.2082)

# ── Story buildings / zones for page 5 ───────────────────────────────────────
# Helper: return first row where link_status matches, or NULL if none exist
.story_row <- function(df, status, fallback = NULL) {
  sub <- df[!is.na(df$link_status) & df$link_status == status, ]
  if (nrow(sub) == 0L) return(fallback)
  sub[1L, ]
}

story_clean    <- .story_row(joined, "in_zone",            joined[1L, ])
story_outside  <- .story_row(joined, "outside_any_zone",   joined[nrow(joined), ])
story_straddle <- .story_row(joined, "straddles_boundary", joined[1L, ])
story_orphan   <- if (nrow(orphan_zones) > 0L) orphan_zones[1L, ] else NULL
story_invalid  <- .story_row(joined, "invalid_geometry",   NULL)

# Compute area_ha for zones if not present
if (!"area_ha" %in% names(zones)) {
  zones_area <- zones |>
    sf::st_transform(crs = 31287L) |>
    dplyr::mutate(area_ha = round(as.numeric(sf::st_area(geometry_wkt)) / 10000, 1)) |>
    sf::st_transform(crs = 4326L)
} else {
  zones_area <- zones
}

# n_buildings per zone for the zone table
zone_bldg_counts <- joined |>
  sf::st_drop_geometry() |>
  dplyr::filter(!is.na(zone_id)) |>
  dplyr::count(zone_id, name = "n_buildings")

zone_table_df <- zones_area |>
  sf::st_drop_geometry() |>
  dplyr::left_join(zone_bldg_counts, by = "zone_id") |>
  dplyr::mutate(n_buildings = dplyr::coalesce(as.integer(n_buildings), 0L)) |>
  dplyr::select(
    zone_id, zone_type, legal_status,
    dplyr::any_of("area_ha"),
    n_buildings
  )

# Buildings per district summary for page 2
district_summary <- joined |>
  sf::st_drop_geometry() |>
  dplyr::count(district, name = "n_buildings") |>
  dplyr::left_join(
    joined |>
      sf::st_drop_geometry() |>
      dplyr::filter(!is.na(zone_id)) |>
      dplyr::count(district, name = "n_in_zone"),
    by = "district"
  ) |>
  dplyr::mutate(
    n_in_zone = dplyr::coalesce(as.integer(n_in_zone), 0L),
    pct_in_zone = round(100 * n_in_zone / n_buildings, 1)
  ) |>
  dplyr::arrange(district)

# decade() helper
decade <- function(year) {
  paste0(floor(year / 10) * 10, "s")
}

# ── Map sample for leaflets (cap at 5000 rows for browser performance) ─────────
# Full data is kept in `joined` for tables, summaries, and crosstabs.
# `joined_for_map` is used only for leaflet markers to avoid browser OOM.
MAP_SAMPLE_N <- 5000L
joined_for_map <- if (nrow(joined) > MAP_SAMPLE_N) {
  set.seed(42L)
  joined[sample.int(nrow(joined), MAP_SAMPLE_N), ]
} else {
  joined
}
```

```{=html}
<script>
document.addEventListener('DOMContentLoaded', function() {
  document.querySelectorAll('[data-bs-toggle="tab"]').forEach(function(tab) {
    tab.addEventListener('shown.bs.tab', function() {
      window.dispatchEvent(new Event('resize'));
    });
  });
  setTimeout(function() { window.dispatchEvent(new Event('resize')); }, 500);
  setTimeout(function() { window.dispatchEvent(new Event('resize')); }, 2000);
});
</script>
```

# Source A — Zones

## Row

::: {.card title="Zone Map"}

```{r}
#| label: fig-zone-map

zones_wgs84 <- sf::st_transform(zones, crs = 4326L)

zone_pal <- leaflet::colorFactor(
  palette = c("#e41a1c", "#377eb8", "#984ea3", "#ff7f0e"),
  domain  = zones_wgs84$zone_type,
  na.color = "#aaaaaa"
)

leaflet(data = zones_wgs84, height = "500px") |>
  leaflet::addProviderTiles(leaflet::providers$CartoDB.Positron) |>
  leaflet::addPolygons(
    fillColor   = ~ zone_pal(zone_type),
    fillOpacity = 0.55,
    color       = "#333333",
    weight      = 1.5,
    opacity     = 1,
    popup       = ~ paste0(
      "<strong>Zone: </strong>", zone_id, "<br>",
      "<strong>Type: </strong>", zone_type, "<br>",
      "<strong>Legal status: </strong>", legal_status
    ),
    label       = ~ paste0(zone_id, " (", zone_type, ")")
  ) |>
  leaflet::addLegend(
    position = "bottomright",
    pal      = zone_pal,
    values   = ~ zone_type,
    title    = "Zone type",
    opacity  = 0.9
  ) |>
  leaflet::setView(
    lng  = vienna_center[["lng"]],
    lat  = vienna_center[["lat"]],
    zoom = 12L
  )
```

:::

## Row

::: {.card title="Zone Attributes"}

```{r}
#| label: tbl-zone-attributes

DT::datatable(
  zone_table_df,
  rownames = FALSE,
  filter   = "top",
  colnames = c("Zone ID", "Type", "Legal status", "Area (ha)", "N buildings"),
  options  = list(
    pageLength = 15,
    scrollX    = TRUE,
    order      = list(list(4L, "desc"))
  )
)
```

:::

::: {.card title="Notes"}

### ACD Zones — Klimaschutzgebiete

Vienna's *Energieraumplanung* (Energy Area Planning, ACD) designates geographic
zones where specific heating and energy-system rules apply to buildings.

**Layer:** `ogdwien:ENERGIEPLANERPOGD`
**Provider:** [Vienna Open Government Data WFS](https://data.wien.gv.at/daten/geo)
**Snapshot date:** `r SNAPSHOT_DATE`
**Total zones in this snapshot:** `r n_zones`

Zone types:

| Type | Legal rule |
|------|-----------|
| `district_heating` | Buildings must connect to district heating network |
| `heat_pump` | Heat-pump installations required; fossil fuel banned |
| `fossil_fuel_ban` | Existing fossil-fuel heating must be retired |

Source columns used from the WFS layer: `ERPLABEL` (zone ID),
`ERPTYP` (zone type), `ERPRECHTSTAT` (legal status).

:::

# Source B — Buildings

## Row

::: {.card title="Building Sample (first 200)"}

```{r}
#| label: fig-building-sample

bldg_sample <- if (nrow(buildings) > 200L) buildings[seq_len(200L), ] else buildings

leaflet(data = bldg_sample, height = "500px") |>
  leaflet::addProviderTiles(leaflet::providers$CartoDB.Positron) |>
  leaflet::addCircleMarkers(
    radius      = 4,
    fillColor   = ~ pal_year(year_built),
    fillOpacity = 0.8,
    color       = "#333333",
    weight      = 0.5,
    opacity     = 0.6,
    popup       = ~ paste0(
      "<strong>Address: </strong>", address, "<br>",
      "<strong>Year built: </strong>", year_built, "<br>",
      "<strong>Height: </strong>", height_m, " m"
    )
  ) |>
  leaflet::addLegend(
    position = "bottomright",
    pal      = pal_year,
    values   = ~ year_built,
    title    = "Year built",
    opacity  = 0.9
  ) |>
  leaflet::setView(
    lng  = vienna_center[["lng"]],
    lat  = vienna_center[["lat"]],
    zoom = 12L
  )
```

:::

## Row

::: {.card title="Buildings by District"}

```{r}
#| label: tbl-district-summary

district_summary |>
  gt::gt() |>
  gt::cols_label(
    district    = "District",
    n_buildings = "Total",
    n_in_zone   = "In a zone",
    pct_in_zone = "% in zone"
  ) |>
  gt::fmt_number(columns = c(n_buildings, n_in_zone), decimals = 0) |>
  gt::fmt_number(columns = pct_in_zone, decimals = 1) |>
  gt::data_color(
    columns = pct_in_zone,
    palette = c("#fee0d2", "#de2d26")
  ) |>
  gt::tab_caption(
    caption = paste0(
      "Buildings per district as of snapshot ", SNAPSHOT_DATE,
      ". 'In a zone' counts buildings with link_status != outside_any_zone."
    )
  )
```

:::

::: {.card title="Year-Built Distribution"}

```{r}
#| label: fig-year-hist

yb <- buildings$year_built[!is.na(buildings$year_built)]
hist(
  yb,
  breaks = 30,
  col    = "#377eb8",
  border = "#1a5276",
  xlab   = "Year built",
  ylab   = "Number of buildings",
  main   = paste0(
    "Year-Built Distribution (N = ", length(yb), ", snapshot ", SNAPSHOT_DATE, ")"
  ),
  las    = 1
)
```

:::

::: {.card title="Notes"}

### Vienna Building Inventory

**Layer:** `ogdwien:GEBAEUDEOGD`
**Provider:** [Vienna Open Government Data WFS](https://data.wien.gv.at/daten/geo)
**Snapshot date:** `r SNAPSHOT_DATE`
**Buildings in this snapshot:** `r format(n_total, big.mark = ",")` (full WFS layer, `max_features = NULL`)

For rendering speed the map above shows the first
`r min(nrow(buildings), 200L)` buildings only.

Column mapping from WFS: `OBJECTID` → `building_id`, `BAUJAHR` → `year_built`,
`HOEHE` / `GEBHOEHE` → `height_m`, `ADRESSE` → `address`.

:::

# Joined: Buildings × Zones

## Row

::: {.card title="Joined Map"}

```{r}
#| label: fig-joined-map

zones_wgs84_join <- sf::st_transform(zones, crs = 4326L)

leaflet(height = "500px") |>
  leaflet::addProviderTiles(leaflet::providers$CartoDB.Positron) |>
  leaflet::addPolygons(
    data        = zones_wgs84_join,
    fillColor   = ~ zone_pal(zone_type),
    fillOpacity = 0.25,
    color       = "#333333",
    weight      = 1.2,
    opacity     = 0.7,
    group       = "Zones",
    popup       = ~ paste0("<strong>Zone: </strong>", zone_id)
  ) |>
  leaflet::addCircleMarkers(
    data        = joined_for_map,
    radius      = 4,
    fillColor   = ~ pal_link(link_status),
    fillOpacity = 0.85,
    color       = "#222222",
    weight      = 0.4,
    opacity     = 0.7,
    group       = "Buildings",
    popup       = ~ paste0(
      "<strong>Building: </strong>", building_id, "<br>",
      "<strong>Status: </strong>", link_status, "<br>",
      "<strong>Zone: </strong>", dplyr::coalesce(zone_id, "none")
    )
  ) |>
  leaflet::addLegend(
    position = "bottomright",
    pal      = pal_link,
    values   = joined$link_status,
    title    = "Link status",
    opacity  = 0.9
  ) |>
  leaflet::addLayersControl(
    overlayGroups = c("Buildings", "Zones"),
    options       = leaflet::layersControlOptions(collapsed = FALSE)
  ) |>
  leaflet::setView(
    lng  = vienna_center[["lng"]],
    lat  = vienna_center[["lat"]],
    zoom = 12L
  )
```

:::

## Row

::: {.card title="Buildings by Decade"}

```{r}
#| label: tbl-decade-marginals

decade_marginals <- joined |>
  sf::st_drop_geometry() |>
  dplyr::filter(!is.na(year_built)) |>
  dplyr::mutate(decade = decade(year_built)) |>
  dplyr::count(decade, name = "total_buildings") |>
  dplyr::left_join(
    joined |>
      sf::st_drop_geometry() |>
      dplyr::filter(!is.na(year_built), !is.na(zone_id)) |>
      dplyr::mutate(decade = decade(year_built)) |>
      dplyr::count(decade, name = "n_in_zone"),
    by = "decade"
  ) |>
  dplyr::mutate(
    n_in_zone   = dplyr::coalesce(as.integer(n_in_zone), 0L),
    pct_in_zone = round(100 * n_in_zone / total_buildings, 1)
  ) |>
  dplyr::arrange(decade) |>
  dplyr::filter(total_buildings > 0L)

decade_marginals |>
  gt::gt() |>
  gt::cols_label(
    decade          = "Decade",
    total_buildings = "Total buildings",
    n_in_zone       = "In a zone",
    pct_in_zone     = "% in zone"
  ) |>
  gt::fmt_number(columns = c(total_buildings, n_in_zone), decimals = 0) |>
  gt::fmt_number(columns = pct_in_zone, decimals = 1) |>
  gt::data_color(
    columns = pct_in_zone,
    palette = c("#fee0d2", "#de2d26")
  ) |>
  gt::tab_caption(
    caption = paste0(
      "Buildings per construction decade and zone coverage. ",
      "N = ", n_total, " buildings, snapshot ", SNAPSHOT_DATE, ". ",
      "Pct in zone = buildings with any zone assignment (link_status != outside_any_zone)."
    )
  )
```

:::

::: {.card title="Link Status Distribution"}

```{r}
#| label: tbl-link-summary

link_summary |>
  dplyr::mutate(
    color_hex = link_status_colors[as.character(link_status)],
    label     = link_status_labels[as.character(link_status)]
  ) |>
  dplyr::select(label, n_buildings, pct) |>
  gt::gt() |>
  gt::cols_label(
    label       = "Link status",
    n_buildings = "N buildings",
    pct         = "% of total"
  ) |>
  gt::fmt_number(columns = n_buildings, decimals = 0) |>
  gt::fmt_number(columns = pct, decimals = 1) |>
  gt::data_color(
    columns = n_buildings,
    palette = c("#e5f5e0", "#31a354")
  ) |>
  gt::tab_caption(
    caption = paste0(
      "Distribution of building-zone linkage outcomes. ",
      "Snapshot: ", SNAPSHOT_DATE, ". Total buildings: ", n_total, "."
    )
  )
```

:::

# Missing Linkage Handling

## Row

::: {.panel-tabset}

### Clean join (in_zone)

```{r}
#| label: fig-case-in-zone

sub <- case_in_zone(joined, n = 50L)

leaflet(data = sub, height = "420px") |>
  leaflet::addProviderTiles(leaflet::providers$CartoDB.Positron) |>
  leaflet::addCircleMarkers(
    radius      = 5,
    fillColor   = link_status_colors[["in_zone"]],
    fillOpacity = 0.85,
    color       = "#1a5276",
    weight      = 0.8,
    popup       = ~ paste0(
      "<strong>Building: </strong>", building_id, "<br>",
      "<strong>Zone: </strong>", zone_id, "<br>",
      "<strong>Overlap fraction: </strong>",
      round(overlap_fraction, 3)
    )
  ) |>
  leaflet::setView(
    lng  = vienna_center[["lng"]],
    lat  = vienna_center[["lat"]],
    zoom = 12L
  )
```

**<span class='good'>HEALTHY</span> — Clean join: building footprint fully
contained in a single zone polygon.**

The building centroid falls inside exactly one zone, and the overlap fraction
is >= 0.95 (at least 95% of the footprint is covered). This is the most
common, unambiguous case — the link is legally unequivocal and the downstream
rule lookup is trivial.

N (this snapshot): **`r n_in_zone`** buildings (`r pct_in_zone`% of total).

### Outside any zone (legitimate)

```{r}
#| label: fig-case-outside

sub_out <- case_outside(joined, n = 100L)

leaflet(data = sub_out, height = "420px") |>
  leaflet::addProviderTiles(leaflet::providers$CartoDB.Positron) |>
  leaflet::addCircleMarkers(
    radius      = 4,
    fillColor   = link_status_colors[["outside_any_zone"]],
    fillOpacity = 0.80,
    color       = "#555555",
    weight      = 0.5,
    popup       = ~ paste0(
      "<strong>Building: </strong>", building_id, "<br>",
      "<strong>District: </strong>", dplyr::coalesce(district, "—"), "<br>",
      "<strong>Year built: </strong>", year_built
    )
  ) |>
  leaflet::setView(
    lng  = vienna_center[["lng"]],
    lat  = vienna_center[["lat"]],
    zoom = 11L
  )
```

**<span class='warn'>INFORMATIONAL</span> — Building outside any ACD zone.**

The building footprint is valid, geocoding succeeded, and the building is in
the official inventory — it simply has no ACD zone overlay at this location.
**This is information, not corruption.** About `r pct_outside`% of Vienna
buildings fall here.

Distinguish from missing data: the building record is complete; the absence
of a zone merely means no ACD energy-planning rule applies to this address.
Common on Vienna's outer districts and undeveloped fringes.

N (this snapshot): **`r n_outside`** buildings (`r pct_outside`% of total).

### Straddles boundary

```{r}
#| label: fig-case-straddle

sub_str <- case_straddle(joined, n = 50L)

leaflet(data = sub_str, height = "420px") |>
  leaflet::addProviderTiles(leaflet::providers$CartoDB.Positron) |>
  leaflet::addCircleMarkers(
    radius      = 5,
    fillColor   = link_status_colors[["straddles_boundary"]],
    fillOpacity = 0.85,
    color       = "#8B4513",
    weight      = 0.8,
    popup       = ~ paste0(
      "<strong>Building: </strong>", building_id, "<br>",
      "<strong>Zone: </strong>", dplyr::coalesce(zone_id, "—"), "<br>",
      "<strong>Overlap fraction: </strong>",
      round(overlap_fraction, 3)
    )
  ) |>
  leaflet::setView(
    lng  = vienna_center[["lng"]],
    lat  = vienna_center[["lat"]],
    zoom = 12L
  )
```

**<span class='warn'>AMBIGUOUS</span> — Building footprint spans two or more
zones (overlap fraction between 5% and 95%).**

The POC uses centroid-based assignment: the building is attributed to whichever
zone contains its centroid. However, legal applicability may cover both zones
for different parts of the building.

**Production fix recommended:** Replace the single `zone_id` with a
list-column of all overlapping zones plus their respective overlap fractions,
and escalate ambiguous cases (overlap fraction 20%–80%) to manual review.

Sample size (this snapshot): **`r n_straddle`** buildings
(`r pct_straddle`% of total).

### Multiple overlapping zones

```{r}
#| label: fig-case-multi

sub_mul <- case_multi(joined, n = 50L)

leaflet(data = sub_mul, height = "420px") |>
  leaflet::addProviderTiles(leaflet::providers$CartoDB.Positron) |>
  leaflet::addCircleMarkers(
    radius      = 5,
    fillColor   = link_status_colors[["multiple_zones"]],
    fillOpacity = 0.85,
    color       = "#4B0082",
    weight      = 0.8,
    popup       = ~ paste0(
      "<strong>Building: </strong>", building_id, "<br>",
      "<strong>Primary zone: </strong>",
      dplyr::coalesce(zone_id, "—"), "<br>",
      "<strong>N overlapping zones: </strong>", n_overlapping_zones
    )
  ) |>
  leaflet::setView(
    lng  = vienna_center[["lng"]],
    lat  = vienna_center[["lat"]],
    zoom = 12L
  )
```

**<span class='warn'>INSPECT</span> — Building centroid falls inside two or
more zone polygons simultaneously.**

Two possible explanations:

1. **Legitimate layering**: Vienna may designate the same area under multiple
   rule frameworks simultaneously (e.g., district-heating zone AND a
   fossil-fuel ban zone covering the same geometry). In this case both rules
   apply and must be stored.

2. **Data error**: Zone polygons that should be mutually exclusive may have
   overlapping geometries due to digitisation errors. Inspect each case.

The POC reports only the first matched zone. Production must store all
overlapping zones as a list-column.

N (this snapshot): **`r n_multi`** buildings (`r pct_multi`% of total).

### Orphan zones (no buildings)

```{r}
#| label: fig-case-orphan

if (nrow(orphan_zones) > 0L) {
  leaflet(data = sf::st_transform(orphan_zones, crs = 4326L), height = "420px") |>
    leaflet::addProviderTiles(leaflet::providers$CartoDB.Positron) |>
    leaflet::addPolygons(
      fillColor   = link_status_colors[["orphan_zone"]],
      fillOpacity = 0.35,
      color       = link_status_colors[["orphan_zone"]],
      weight      = 2.0,
      opacity     = 0.9,
      popup       = ~ paste0(
        "<strong>Zone: </strong>", zone_id, "<br>",
        "<strong>Type: </strong>", zone_type
      )
    ) |>
    leaflet::setView(
      lng  = vienna_center[["lng"]],
      lat  = vienna_center[["lat"]],
      zoom = 11L
    )
} else {
  cat("No orphan zones in this snapshot — all zones contain at least one building.")
}
```

**<span class='warn'>INVESTIGATE</span> — Zones containing zero building
centroids.**

N orphan zones (this snapshot): **`r n_orphan`**.

Possible causes:

- **(a) Newly designated zone over undeveloped land** — legitimate. The zone
  designates future development requirements; no current buildings to classify.
- **(b) Inventory snapshot lag** — buildings exist physically but are not yet
  catalogued in the `GEBAEUDEOGD` inventory. The zone polygon is correct; the
  building list is incomplete.
- **(c) Zone polygon digitisation error** — the polygon was drawn in the wrong
  location. Inspect against imagery.

Investigate each orphan individually before concluding it is a data error.

### Invalid geometry

```{r}
#| label: fig-case-invalid

sub_inv <- case_invalid(joined)

if (nrow(sub_inv) > 0L) {
  leaflet(data = sub_inv, height = "420px") |>
    leaflet::addProviderTiles(leaflet::providers$CartoDB.Positron) |>
    leaflet::addCircleMarkers(
      radius      = 6,
      fillColor   = link_status_colors[["invalid_geometry"]],
      fillOpacity = 0.90,
      color       = "#8B0000",
      weight      = 1.0,
      popup       = ~ paste0(
        "<strong>Building: </strong>", building_id, "<br>",
        "<strong>Address: </strong>", address
      )
    ) |>
    leaflet::setView(
      lng  = vienna_center[["lng"]],
      lat  = vienna_center[["lat"]],
      zoom = 12L
    )
} else {
  cat(paste0(
    "No invalid-geometry buildings in this snapshot (N = 0). ",
    "st_make_valid() was applied during fetch — all footprints passed validation."
  ))
}
```

**`r if (n_invalid == 0L) "<span class='good'>PASS</span>" else "<span class='bad'>FAIL</span>"` — Self-intersecting or otherwise broken polygons.**

N invalid buildings (this snapshot): **`r n_invalid`**.

The join pipeline calls `sf::st_make_valid()` on all input geometries before
performing any spatial operations. If this count is non-zero after that step,
document the upstream data issue and report it to the data provider.

Expected in a clean run: **0**.

### CRS mismatch

**No map for this case** — CRS errors are caught at code time, not at render time.

**CRS chain used in this POC:**

| Stage | CRS | EPSG | Purpose |
|-------|-----|------|---------|
| Native WFS data | MGI / Austria GK East | 31256 | Source CRS from data provider |
| Working (analysis) | MGI / Austria Lambert | 31287 | Area + distance calculations |
| Display (leaflet) | WGS84 | 4326 | Leaflet rendering |

A CRS mismatch between buildings and zones causes **silent ~100 m offsets** in
centroid-to-zone intersection — buildings that should be inside a zone appear
outside it and vice versa.

**Protection:** `join_buildings_zones()` contains an explicit assert that both
inputs are in EPSG:31287. If either input differs, the function stops with an
informative error before any spatial operation runs:

```r
if (b_crs != CRS_WORKING || z_crs != CRS_WORKING) {
  cli::cli_abort(
    c("Both inputs must be in EPSG:{CRS_WORKING}.",
      "i" = "{.arg buildings} CRS: EPSG:{b_crs}",
      "i" = "{.arg zones} CRS: EPSG:{z_crs}")
  )
}
```

### Snapshot lag

**No map for this case** — temporal coverage is a data provenance issue,
not a spatial geometry issue.

**The pinned snapshot is the contract.**

Buildings constructed **after** `r SNAPSHOT_DATE` cannot appear in this
analysis regardless of their spatial location. The snapshot date is defined
once in `R/snapshot.R`:

```r
SNAPSHOT_DATE <- "`r SNAPSHOT_DATE`"
```

To refresh the analysis:

1. Bump `SNAPSHOT_DATE` in `R/snapshot.R`.
2. Re-run `fetch_acd_zones()` and `fetch_buildings()` — these write new
   Parquet files to `inst/extdata/zones_<date>.parquet` and
   `inst/extdata/buildings_<date>.parquet`.
3. Re-run `join_buildings_zones()` to produce the new joined Parquet.
4. Re-render this dashboard.

Do not compare analysis results across snapshots without accounting for the
change in building inventory.

### ACD agreement (cross-validation)

::: {.card title="Confusion matrix"}

```{r}
#| label: tbl-acd-confusion-matrix

confusion_matrix_df |>
  gt::gt(rowname_col = "Spatial join result") |>
  gt::tab_header(
    title    = "Spatial join vs City ACD cross-validation",
    subtitle = paste0("Snapshot: ", SNAPSHOT_DATE, " — N = ", n_total, " buildings")
  ) |>
  gt::tab_spanner(
    label   = "City of Vienna ACD field",
    columns = dplyr::starts_with("City ACD")
  ) |>
  gt::tab_caption(
    caption = paste0(
      "2×2 cross-validation of spatial join (rows) vs city ACD field (columns). ",
      "Snapshot: ", SNAPSHOT_DATE, ". Total buildings: ", n_total, ". ",
      "Agreement: ", n_agree_total, " buildings; Disagreements: ", n_disagree_total, "."
    )
  )
```

:::

::: {.card title="Disagreements on the map"}

```{r}
#| label: fig-acd-disagreement-map

if (nrow(acd_disagree_sf) > 0L) {
  leaflet(data = acd_disagree_sf, height = "420px") |>
    leaflet::addProviderTiles(leaflet::providers$CartoDB.Positron) |>
    leaflet::addCircleMarkers(
      radius      = 5,
      fillColor   = ~ pal_disagree(as.character(acd_agreement)),
      fillOpacity = 0.85,
      color       = "#333333",
      weight      = 0.6,
      opacity     = 0.7,
      group       = ~ as.character(acd_agreement),
      popup       = ~ paste0(
        "<strong>Building: </strong>", building_id, "<br>",
        "<strong>Agreement: </strong>", acd_agreement, "<br>",
        "<strong>City ACD: </strong>", dplyr::coalesce(acd_native, "(empty)"), "<br>",
        "<strong>Spatial join zone: </strong>", dplyr::coalesce(zone_id, "(none)")
      )
    ) |>
    leaflet::addLegend(
      position = "bottomright",
      pal      = pal_disagree,
      values   = ~ as.character(acd_agreement),
      title    = "Disagreement type",
      opacity  = 0.9
    ) |>
    leaflet::addLayersControl(
      overlayGroups = c("spatial_only", "acd_only"),
      options       = leaflet::layersControlOptions(collapsed = FALSE)
    ) |>
    leaflet::setView(
      lng  = vienna_center[["lng"]],
      lat  = vienna_center[["lat"]],
      zoom = 12L
    )
} else {
  cat("No disagreements found in this snapshot — spatial join and city ACD agree on all buildings.")
}
```

:::

::: {.card title="Notes"}

### What this cross-validation tests

The City of Vienna already tags every building with an ACD identifier in
the `ogdwien:GEBAEUDEINFOOGD` layer (the `ACD` field). Our spatial join is
an **independent computation** — we overlay building footprints onto zone
polygons geometrically, without consulting the city's pre-linked field.

Disagreements are **interesting findings**, not data errors:

- **`spatial_only`** (`r n_spatial_only` buildings): our spatial join finds
  a zone, but the city's `ACD` field is empty or `NA`. Possible under-coverage
  in the city's pre-linking step.
- **`acd_only`** (`r n_acd_only` buildings): the city's `ACD` field is
  populated, but our spatial join finds no overlapping zone. Possible stale
  zone geometry, or a zone that was present when the city's field was computed
  but has since been modified in the WFS layer.

Total agreements in this snapshot: **`r n_agree_total`** buildings.
Total disagreements: **`r n_disagree_total`** buildings.

**Coverage:** This snapshot is the full `ogdwien:GEBAEUDEINFOOGD` WFS
layer — `r format(n_total, big.mark = ",")` buildings. Any `acd_only` count
in this view reflects a real divergence between the city's pre-linked `ACD`
field and the independent spatial join, not a sampling artefact.

:::

:::

# Stories — Worked Well / Failed Badly

## Row

::: {.panel-tabset}

### Worked

::: {.panel-tabset}

#### Story 1: The cleanly classified apartment block

```{r}
#| label: fig-story-clean

story_lat <- sf::st_coordinates(sf::st_centroid(story_clean))[, 2]
story_lng <- sf::st_coordinates(sf::st_centroid(story_clean))[, 1]

# Find the zone this building belongs to
story_zone_id <- story_clean$zone_id[[1]]
story_zone_sf <- zones[!is.na(zones$zone_id) & zones$zone_id == story_zone_id, ]

m <- leaflet(height = "400px") |>
  leaflet::addProviderTiles(leaflet::providers$CartoDB.Positron)

if (nrow(story_zone_sf) > 0L) {
  m <- m |>
    leaflet::addPolygons(
      data        = story_zone_sf,
      fillColor   = "#e41a1c",
      fillOpacity = 0.30,
      color       = "#e41a1c",
      weight      = 2.0,
      popup       = ~ paste0("Zone: ", zone_id)
    )
}

m |>
  leaflet::addCircleMarkers(
    data        = story_clean,
    radius      = 10,
    fillColor   = link_status_colors[["in_zone"]],
    fillOpacity = 0.95,
    color       = "#1a5276",
    weight      = 2.0,
    popup       = ~ paste0(
      "<strong>Building: </strong>", building_id, "<br>",
      "<strong>Zone: </strong>", dplyr::coalesce(zone_id, "—"), "<br>",
      "<strong>Year built: </strong>", year_built, "<br>",
      "<strong>Overlap fraction: </strong>",
      round(overlap_fraction, 3)
    )
  ) |>
  leaflet::setView(lng = story_lng, lat = story_lat, zoom = 15L)
```

Building **`r story_clean$building_id[[1]]`**, constructed
**`r story_clean$year_built[[1]]`**, falls entirely within Zone
**`r dplyr::coalesce(story_clean$zone_id[[1]], "(none)")`**
(`r dplyr::coalesce(story_clean$zone_type[[1]], "unknown")` — legal status:
`r dplyr::coalesce(story_clean$legal_status[[1]], "unknown")`).

The centroid-based join assigns this building unambiguously to one zone.
Overlap fraction: `r round(story_clean$overlap_fraction[[1]], 3)`. No manual
review required. Downstream the legal rule lookup is a simple table join on
`zone_id`.

#### Story 2: Zone-free outskirts

```{r}
#| label: fig-story-outside

outside_lat <- sf::st_coordinates(sf::st_centroid(story_outside))[, 2]
outside_lng <- sf::st_coordinates(sf::st_centroid(story_outside))[, 1]

leaflet(height = "400px") |>
  leaflet::addProviderTiles(leaflet::providers$CartoDB.Positron) |>
  leaflet::addCircleMarkers(
    data        = story_outside,
    radius      = 10,
    fillColor   = link_status_colors[["outside_any_zone"]],
    fillOpacity = 0.90,
    color       = "#555555",
    weight      = 1.5,
    popup       = ~ paste0(
      "<strong>Building: </strong>", building_id, "<br>",
      "<strong>District: </strong>",
      dplyr::coalesce(district, "—"), "<br>",
      "<strong>Year built: </strong>", year_built
    )
  ) |>
  leaflet::setView(lng = outside_lng, lat = outside_lat, zoom = 14L)
```

Building **`r story_outside$building_id[[1]]`** sits in district
**`r dplyr::coalesce(story_outside$district[[1]], "unknown")`** with no ACD
zone designation. The building owner is currently free to install any heating
system — no ACD energy-planning rule binds this address as of
`r SNAPSHOT_DATE`.

This is not a data error. The building record is complete (year built:
`r story_outside$year_built[[1]]`, height: `r story_outside$height_m[[1]]` m,
address: `r story_outside$address[[1]]`). The absence of a zone assignment is
a valid, informative outcome. It means: *no ACD action required for this
property*.

:::

### Failed

::: {.panel-tabset}

#### Story 3: The straddler — half a building, half a rule

```{r}
#| label: fig-story-straddle

straddle_lat <- sf::st_coordinates(sf::st_centroid(story_straddle))[, 2]
straddle_lng <- sf::st_coordinates(sf::st_centroid(story_straddle))[, 1]

leaflet(height = "400px") |>
  leaflet::addProviderTiles(leaflet::providers$CartoDB.Positron) |>
  leaflet::addPolygons(
    data        = zones,
    fillColor   = ~ zone_pal(zone_type),
    fillOpacity = 0.30,
    color       = "#333333",
    weight      = 1.5,
    popup       = ~ paste0("Zone: ", zone_id)
  ) |>
  leaflet::addCircleMarkers(
    data        = story_straddle,
    radius      = 10,
    fillColor   = link_status_colors[["straddles_boundary"]],
    fillOpacity = 0.95,
    color       = "#8B4513",
    weight      = 2.0,
    popup       = ~ paste0(
      "<strong>Building: </strong>", building_id, "<br>",
      "<strong>Primary zone: </strong>",
      dplyr::coalesce(zone_id, "—"), "<br>",
      "<strong>Overlap fraction: </strong>",
      round(overlap_fraction, 3), "<br>",
      "<strong>N overlapping zones: </strong>", n_overlapping_zones
    )
  ) |>
  leaflet::setView(lng = straddle_lng, lat = straddle_lat, zoom = 15L)
```

Building **`r story_straddle$building_id[[1]]`** sits with
**`r round(story_straddle$overlap_fraction[[1]] * 100, 0)`%** of its footprint
inside zone **`r dplyr::coalesce(story_straddle$zone_id[[1]], "(none)")`**.
The remaining
**`r round((1 - story_straddle$overlap_fraction[[1]]) * 100, 0)`%** lies
outside that zone — potentially inside another zone.

**POC limitation:** The centroid-based join assigns this building to the zone
containing its centroid. The reported `zone_id` is
**`r dplyr::coalesce(story_straddle$zone_id[[1]], "none")`** and the reported
`link_status` is `straddles_boundary` as a warning signal.

**Production fix:** Store a list-column of all overlapping zones with
individual overlap fractions. Legal teams should handle boundary-straddling
buildings case-by-case.

#### Story 4: The orphan polygon

```{r}
#| label: fig-story-orphan

if (!is.null(story_orphan) && nrow(story_orphan) > 0L) {
  orphan_lat <- sf::st_coordinates(sf::st_centroid(story_orphan))[, 2]
  orphan_lng <- sf::st_coordinates(sf::st_centroid(story_orphan))[, 1]

  leaflet(height = "400px") |>
    leaflet::addProviderTiles(leaflet::providers$CartoDB.Positron) |>
    leaflet::addPolygons(
      data        = story_orphan,
      fillColor   = link_status_colors[["orphan_zone"]],
      fillOpacity = 0.40,
      color       = link_status_colors[["orphan_zone"]],
      weight      = 2.5,
      popup       = ~ paste0(
        "<strong>Zone: </strong>", zone_id, "<br>",
        "<strong>Type: </strong>", zone_type, "<br>",
        "<strong>Buildings inside: </strong>0"
      )
    ) |>
    leaflet::setView(lng = orphan_lng, lat = orphan_lat, zoom = 14L)
} else {
  cat("No orphan zones in this snapshot.")
}
```

```{r}
#| label: prose-story-orphan
#| output: asis

if (!is.null(story_orphan) && nrow(story_orphan) > 0L) {
  orphan_area <- tryCatch(
    round(
      as.numeric(sf::st_area(sf::st_transform(story_orphan, crs = 31287L))) / 10000,
      1
    ),
    error = function(e) NA_real_
  )
  cat(paste0(
    "Zone **", story_orphan$zone_id[[1]], "**",
    " (`", story_orphan$zone_type[[1]], "`)",
    " covers",
    if (!is.na(orphan_area)) paste0(" approximately **", orphan_area, " ha**") else " an unknown area",
    " but contains **zero building centroids** in the `", SNAPSHOT_DATE, "` snapshot.\n\n",
    "This could be: (a) a newly designated zone over undeveloped land —",
    " **legitimate**; (b) an inventory lag; or (c) a polygon digitisation error.",
    " Compare against Vienna satellite imagery to distinguish the cases.",
    "\n\nOrphan zones should not be conflated with the `outside_any_zone`",
    " building classification: the zone is real — it is the building list",
    " that may be incomplete relative to the zone."
  ))
} else {
  cat(paste0(
    "No orphan zones in the `", SNAPSHOT_DATE, "` snapshot — all zones contain",
    " at least one building centroid. This is a healthy result."
  ))
}
```

#### Story 5: The case of the impossible building

```{r}
#| label: prose-story-invalid
#| output: asis

if (!is.null(story_invalid) && nrow(story_invalid) > 0L) {
  cat(paste0(
    "Building **", story_invalid$building_id[[1]], "**",
    " at address **", story_invalid$address[[1]], "**",
    " has a self-intersecting footprint polygon as delivered by the WFS.\n\n",
    "After calling `sf::st_make_valid()` the geometry is repaired and the",
    " building participates in the join normally. However, we flag it as",
    " `invalid_geometry` in the output so that the data quality dashboard",
    " surfaces the upstream issue.\n\n",
    "**Why this matters:** an undetected invalid geometry would cause",
    " `sf::st_intersects()` to return unpredictable results and silently",
    " exclude the building from zone assignment. Documenting it here creates",
    " an audit trail for the data provider.\n\n",
    "N invalid in this snapshot: **", n_invalid, "**."
  ))
} else {
  cat(paste0(
    "No invalid-geometry buildings in the `", SNAPSHOT_DATE, "` snapshot.",
    " `sf::st_make_valid()` resolved all geometries before the join.",
    " This story would fire if any upstream footprint polygon is",
    " self-intersecting or otherwise malformed."
  ))
}
```

:::

:::

# Join QA

## Row

::: {.card title="Link Status Distribution"}

```{r}
#| label: tbl-qa-link-dist

link_summary |>
  dplyr::mutate(
    status_label = link_status_labels[as.character(link_status)],
    color_hex    = link_status_colors[as.character(link_status)]
  ) |>
  dplyr::select(status_label, n_buildings, pct, color_hex) |>
  gt::gt() |>
  gt::cols_label(
    status_label = "Link status",
    n_buildings  = "N buildings",
    pct          = "%",
    color_hex    = "Color"
  ) |>
  gt::fmt_number(columns = n_buildings, decimals = 0) |>
  gt::fmt_number(columns = pct, decimals = 1) |>
  gt::data_color(
    columns = n_buildings,
    palette = c("#f7fbff", "#08306b")
  ) |>
  gt::cols_width(color_hex ~ gt::px(60)) |>
  gt::tab_caption(
    caption = paste0(
      "Full link-status distribution. Snapshot: ", SNAPSHOT_DATE,
      ". Total buildings: ", n_total, "."
    )
  )
```

:::

::: {.card title="Coverage Stats"}

**Snapshot:** `r SNAPSHOT_DATE`

| Metric | Value |
|--------|-------|
| Total buildings | `r n_total` |
| Total zones | `r n_zones` |
| Buildings in any zone | `r n_total - n_outside` (`r 100 - pct_outside`%) |
| Buildings outside all zones | `r n_outside` (`r pct_outside`%) |
| Buildings straddling boundary | `r n_straddle` (`r pct_straddle`%) |
| Buildings under multiple zones | `r n_multi` (`r pct_multi`%) |
| Orphan zones (no buildings) | `r n_orphan` |
| Invalid-geometry buildings | `r n_invalid` |
| Using placeholder data | `r using_placeholder` |

:::

## Row

::: {.card title="Sanity Checks"}

```{r}
#| label: tbl-sanity-checks

checks <- tibble::tibble(
  check = c(
    "All buildings have valid geometry (st_is_valid = TRUE)",
    "All zones have valid geometry (st_is_valid = TRUE)",
    "Sum of link_status percentages = 100%",
    "No NA building_id values",
    "link_status factor levels complete"
  ),
  result = c(
    if (n_invalid == 0L) "PASS" else paste0("FAIL — ", n_invalid, " invalid"),
    tryCatch({
      n_bad_z <- sum(!sf::st_is_valid(zones), na.rm = TRUE)
      if (n_bad_z == 0L) "PASS" else paste0("FAIL — ", n_bad_z, " invalid zones")
    }, error = function(e) "WARN — could not check"),
    if (abs(sum(link_summary$pct) - 100) < 0.2) "PASS" else
      paste0("FAIL — sum = ", round(sum(link_summary$pct), 1), "%"),
    if (sum(is.na(joined$building_id)) == 0L) "PASS" else
      paste0("FAIL — ", sum(is.na(joined$building_id)), " NAs"),
    if (all(
      c("in_zone", "outside_any_zone", "straddles_boundary",
        "multiple_zones", "orphan_zone", "invalid_geometry") %in%
      levels(joined$link_status)
    )) "PASS" else "FAIL — missing levels"
  )
)

checks |>
  gt::gt() |>
  gt::cols_label(check = "Check", result = "Result") |>
  gt::tab_style(
    style     = gt::cell_fill(color = "#d4edda"),
    locations = gt::cells_body(
      columns = result,
      rows    = grepl("^PASS", result)
    )
  ) |>
  gt::tab_style(
    style     = gt::cell_fill(color = "#f8d7da"),
    locations = gt::cells_body(
      columns = result,
      rows    = grepl("^FAIL", result)
    )
  ) |>
  gt::tab_style(
    style     = gt::cell_fill(color = "#fff3cd"),
    locations = gt::cells_body(
      columns = result,
      rows    = grepl("^WARN", result)
    )
  ) |>
  gt::tab_caption(caption = paste0(
    "Automated sanity checks on joined data. Snapshot: ", SNAPSHOT_DATE, "."
  ))
```

:::

::: {.card title="Recommendations"}

### What needs follow-up

1. **Straddles-boundary cases (`r n_straddle` buildings):** The POC assigns
   one zone via centroid. Production should compute area-weighted overlap and
   store a list-column of all applicable zones with fractions.

2. **Multiple-zone cases (`r n_multi` buildings):** Audit each manually to
   determine whether overlapping zone polygons are intentional (layered rules)
   or a digitisation error. If intentional, the data model must support
   multiple simultaneous zone assignments per building.

3. **Orphan zones (`r n_orphan`):** Cross-reference each orphan polygon against
   current satellite imagery and planning documents to classify as
   (a) legitimate undeveloped land, (b) inventory lag, or (c) polygon error.

4. **District assignment:** The `district` column is `NA` for all buildings in
   this POC. Production should perform a secondary spatial join against the
   Vienna district boundary layer (`BEZIRKSGRENZEOGD`) to populate this field.

### What production should add

- **Area-weighted overlap:** Replace centroid join with full footprint
  intersection to compute exact overlap fractions per zone.
- **List-column schema:** Allow multiple `(zone_id, zone_type, overlap_fraction)`
  tuples per building — required for straddlers and multi-zone buildings.
- **Address geocoding fallback:** For buildings with invalid geometry, fall back
  to address-based geocoding before marking as invalid.
- **Snapshot refresh pipeline:** Automate `SNAPSHOT_DATE` bumping and
  re-fetching via a targets pipeline target.
- **pointblank data contract:** Convert these sanity checks into a formal
  pointblank agent with alerting on refresh.

:::

## Row

::: {.card title="Remote duckplyr query (Stage 2 preview)"}

```{r}
#| label: tbl-remote-duckplyr-query

# Stage 2: demonstrate live duckplyr query against Parquet hosted on GitHub Pages.
# DuckDB issues HTTP range requests — only the link_status column is downloaded,
# not the full file. Gracefully degrades if Pages is not yet live or duckplyr
# is not available in the render environment.
demo_remote_query <- tryCatch({
  acd::load_acd_data("joined", source = "remote", collect = FALSE) |>
    dplyr::count(.data$link_status, name = "n") |>
    dplyr::collect()
}, error = function(e) {
  message("Remote duckplyr query unavailable: ", conditionMessage(e))
  NULL
})

if (!is.null(demo_remote_query)) {
  demo_remote_query |>
    gt::gt() |>
    gt::cols_label(link_status = "Link status", n = "N buildings") |>
    gt::fmt_number(columns = n, decimals = 0) |>
    gt::tab_caption(
      caption = paste0(
        "Live duckplyr query against Parquet served by GitHub Pages. ",
        "DuckDB issues HTTP range requests; only the link_status column is downloaded, ",
        "not the full file. Source URL: ",
        "https://johngavin.github.io/acd_area_climate_design/data/joined_",
        SNAPSHOT_DATE, ".parquet"
      )
    )
} else {
  cat(paste0(
    "Remote query unavailable: GitHub Pages may not yet be live, ",
    "or duckplyr is not available in this render environment. ",
    "The local snapshot (N = ", n_total, " buildings) was used throughout the rest of this dashboard."
  ))
}
```

:::

::: {.card title="Stage 2 architecture notes"}

### Parquet on GitHub Pages

The snapshot Parquet files are now committed to `docs/data/` and served by
GitHub Pages at:

`https://johngavin.github.io/acd_area_climate_design/data/`

The `load_acd_data()` function supports two modes:

| `source` | Reads from | Speed | Network |
|----------|-----------|-------|---------|
| `"local"` (default) | `inst/extdata/` (committed) | Instant | None |
| `"remote"` | GitHub Pages via DuckDB httpfs | ~1–2 s | HTTP range requests |

**Why HTTP range requests are efficient:** DuckDB reads only the byte ranges
needed for the query (predicate / projection pushdown). A `count(link_status)`
query downloads only the `link_status` column bytes — typically < 1 KB from a
30 KB Parquet file.

**Stage 3 upgrade path:** Replace `REMOTE_BASE_URL` in `R/snapshot.R` with an
HuggingFace `hf://datasets/...` or S3/R2 URL when the dataset outgrows GitHub
Pages limits (100 MB file cap, 1 GB repo cap).

:::

