---
params:
  data_source: "local"
format:
  dashboard:
    orientation: rows
    scrolling: true
    theme:
      - cosmo
      - custom.scss
execute:
  echo: false
  message: false
  warning: false
---

```{r}
#| label: setup
#| include: false

library(sf)
library(leaflet)
library(dplyr)
library(tidyr)
library(DT)
library(gt)

# ── Package loading ────────────────────────────────────────────────────────────
# Try pkgload first (dev mode), then library (installed), then graceful fail.
tryCatch(
  pkgload::load_all(quiet = TRUE),
  error = function(e) {
    tryCatch(library(acd), error = function(e2) NULL)
  }
)

# Snapshot date constant (defined in R/snapshot.R; fallback if not loaded)
if (!exists("SNAPSHOT_DATE")) SNAPSHOT_DATE <- "2026-05-01"

# ── Data source selection (Stage 2: "remote" uses GitHub Pages Parquet) ────────
# Default: "local" — reads from inst/extdata (fast, offline, always available).
# Set to "remote" to test the duckplyr path against GitHub Pages Parquet.
# Can also be overridden via params: in YAML front-matter for testing.
data_source <- if (!is.null(params$data_source)) params$data_source else "local"

# ── Parquet cache helper ───────────────────────────────────────────────────────
# Resolves the inst/extdata path for both installed and dev-tree usage,
# then reads via arrow::read_parquet().
.cached_parquet <- function(filename) {
  installed <- system.file("extdata", filename, package = "acd")
  if (nzchar(installed) && file.exists(installed)) {
    return(arrow::read_parquet(installed))
  }
  dev_path <- tryCatch(
    file.path(find.package("acd"), "inst", "extdata", filename),
    error = function(e) ""
  )
  if (nzchar(dev_path) && file.exists(dev_path)) {
    return(arrow::read_parquet(dev_path))
  }
  vignette_dir <- tryCatch(
    dirname(normalizePath(knitr::current_input())),
    error = function(e) ""
  )
  if (nzchar(vignette_dir)) {
    src_path <- normalizePath(
      file.path(vignette_dir, "..", "..", "inst", "extdata", filename),
      mustWork = FALSE
    )
    if (file.exists(src_path)) return(arrow::read_parquet(src_path))
  }
  NULL
}

# ── Placeholder data (used when real parquet is absent) ────────────────────────
# Creates a synthetic sf with all 6 link_status values so the dashboard
# renders completely before real WFS data arrives.
.placeholder_joined <- function() {
  set.seed(42L)
  n <- 18L

  # Vienna centre coordinates + small offsets (degrees)
  lat_c <- 48.2082
  lng_c <- 16.3725

  lats <- lat_c + runif(n, -0.03, 0.03)
  lngs <- lng_c + runif(n, -0.04, 0.04)

  # Create point geometries in WGS84 (for placeholder — display only)
  geom <- sf::st_sfc(
    lapply(seq_len(n), function(i) sf::st_point(c(lngs[i], lats[i]))),
    crs = 4326L
  )

  link_statuses <- factor(
    c(
      rep("in_zone",            7L),
      rep("outside_any_zone",   4L),
      rep("straddles_boundary", 3L),
      rep("multiple_zones",     2L),
      rep("orphan_zone",        1L),
      rep("invalid_geometry",   1L)
    ),
    levels = c(
      "in_zone", "outside_any_zone", "straddles_boundary",
      "multiple_zones", "orphan_zone", "invalid_geometry"
    )
  )

  zone_ids <- c(
    rep("ZONE-A1", 4L), rep("ZONE-B2", 3L), rep(NA_character_, 4L),
    rep("ZONE-A1", 3L), rep(c("ZONE-A1", "ZONE-B2"), 1L),
    "ZONE-C3", NA_character_
  )

  zone_types <- dplyr::case_when(
    link_statuses == "outside_any_zone" ~ NA_character_,
    zone_ids == "ZONE-A1" ~ "district_heating",
    zone_ids == "ZONE-B2" ~ "heat_pump",
    zone_ids == "ZONE-C3" ~ "fossil_fuel_ban",
    TRUE ~ NA_character_
  )

  legal_statuses <- dplyr::case_when(
    zone_types == "district_heating" ~ "district_heating_mandatory",
    zone_types == "heat_pump"        ~ "heat_pump_required",
    zone_types == "fossil_fuel_ban"  ~ "fossil_fuel_banned",
    TRUE ~ NA_character_
  )

  district_labels <- c(
    rep("01", 4L), rep("02", 4L), rep("10", 4L),
    rep("13", 3L), rep("22", 3L)
  )

  sf::st_sf(
    building_id         = paste0("B-SYNTH-", sprintf("%03d", seq_len(n))),
    district            = district_labels,
    year_built          = as.integer(sample(1880:2020, n, replace = TRUE)),
    height_m            = round(runif(n, 8, 40), 1),
    address             = paste0("Musterstrasse ", seq_len(n), ", Wien"),
    zone_id             = zone_ids,
    zone_type           = zone_types,
    legal_status        = legal_statuses,
    link_status         = link_statuses,
    n_overlapping_zones = dplyr::case_when(
      link_statuses == "outside_any_zone"   ~ 0L,
      link_statuses == "multiple_zones"     ~ 2L,
      link_statuses == "straddles_boundary" ~ 1L,
      TRUE ~ 1L
    ),
    overlap_fraction    = dplyr::case_when(
      link_statuses == "in_zone"            ~ 0.98,
      link_statuses == "outside_any_zone"   ~ 0.00,
      link_statuses == "straddles_boundary" ~ round(runif(
        sum(link_statuses == "straddles_boundary"), 0.10, 0.90), 2),
      link_statuses == "multiple_zones"     ~ 0.70,
      TRUE ~ 0.00
    ),
    geometry = geom
  )
}

# ── Placeholder zones (3 fake zones as polygons) ───────────────────────────────
.placeholder_zones <- function() {
  lat_c <- 48.2082
  lng_c <- 16.3725

  make_box <- function(lat0, lng0, dlat = 0.012, dlng = 0.016) {
    sf::st_polygon(list(matrix(
      c(lng0,        lat0,
        lng0 + dlng, lat0,
        lng0 + dlng, lat0 + dlat,
        lng0,        lat0 + dlat,
        lng0,        lat0),
      ncol = 2, byrow = TRUE
    )))
  }

  geom <- sf::st_sfc(
    make_box(lat_c - 0.01,  lng_c - 0.02),
    make_box(lat_c + 0.005, lng_c + 0.005),
    make_box(lat_c - 0.025, lng_c + 0.01),
    crs = 4326L
  )

  sf::st_sf(
    zone_id      = c("ZONE-A1", "ZONE-B2", "ZONE-C3"),
    zone_type    = c("district_heating", "heat_pump", "fossil_fuel_ban"),
    legal_status = c(
      "district_heating_mandatory",
      "heat_pump_required",
      "fossil_fuel_banned"
    ),
    area_ha      = c(12.4, 8.7, 5.1),
    geometry     = geom
  )
}

# ── link_status summary helper (inline fallback if acd package not loaded) ─────
.link_status_summary_local <- function(joined) {
  joined |>
    sf::st_drop_geometry() |>
    dplyr::count(link_status, name = "n_buildings") |>
    dplyr::mutate(
      pct = round(100 * n_buildings / sum(n_buildings), 1)
    ) |>
    dplyr::arrange(link_status)
}

# ── Load or synthesise data ────────────────────────────────────────────────────
# Stage 1: try load_acd_data() (package function); fall back to .cached_parquet;
# fall back to placeholder when neither source is available.
joined_raw <- tryCatch(
  {
    df <- acd::load_acd_data("joined", source = data_source)
    df
  },
  error = function(e) {
    message("load_acd_data() failed (", conditionMessage(e), "); trying .cached_parquet fallback")
    .cached_parquet(paste0("joined_", SNAPSHOT_DATE, ".parquet"))
  }
)

using_placeholder <- is.null(joined_raw)

if (using_placeholder) {
  joined <- .placeholder_joined()
} else {
  # Convert to sf if parquet stripped the geometry column
  if (!inherits(joined_raw, "sf")) {
    joined <- sf::st_as_sf(joined_raw, wkt = "geometry_wkt", crs = 31287L)
  } else {
    joined <- joined_raw
  }
  # Transform to WGS84 for leaflet
  joined <- sf::st_transform(joined, crs = 4326L)
  # Add compatibility aliases for columns renamed in the real schema
  if (!"zone_type" %in% names(joined) && "acd_native" %in% names(joined)) {
    joined$zone_type <- as.character(joined$acd_native)
  }
  if (!"legal_status" %in% names(joined)) {
    joined$legal_status <- NA_character_
  }
  if (!"height_m" %in% names(joined) && "n_stories" %in% names(joined)) {
    joined$height_m <- as.numeric(joined$n_stories)
  }
}

# Zones: derive from joined or load separately
zones_raw <- .cached_parquet(paste0("zones_", SNAPSHOT_DATE, ".parquet"))
if (is.null(zones_raw)) {
  zones <- .placeholder_zones()
} else {
  if (!inherits(zones_raw, "sf")) {
    zones <- sf::st_as_sf(zones_raw, wkt = "geometry_wkt", crs = 31287L)
  } else {
    zones <- zones_raw
  }
  # Add canonical zone_id from ERPLABEL if not already present
  if (!"zone_id" %in% names(zones) && "ERPLABEL" %in% names(zones)) {
    zones$zone_id <- zones$ERPLABEL
  }
  # Add zone_type alias from acd_native if present on zones (else use placeholder)
  if (!"zone_type" %in% names(zones)) {
    zones$zone_type <- NA_character_
  }
  # Add legal_status alias
  if (!"legal_status" %in% names(zones)) {
    zones$legal_status <- NA_character_
  }
  zones <- sf::st_transform(zones, crs = 4326L)
}

# Buildings: subset of joined that are actual building points/polygons
buildings <- joined

# ── Pre-compute summary objects ────────────────────────────────────────────────
# link_status summary — use package function if available, else local fallback
link_summary <- tryCatch(
  acd:::link_status_summary(joined),
  error = function(e) .link_status_summary_local(joined)
)

# Counts by link_status for inline prose
n_total      <- nrow(joined)
n_zones      <- nrow(zones)

ls_counts <- link_summary |>
  dplyr::select(link_status, n_buildings) |>
  tibble::deframe()

n_in_zone     <- ls_counts[["in_zone"]]            %||% 0L
n_outside     <- ls_counts[["outside_any_zone"]]   %||% 0L
n_straddle    <- ls_counts[["straddles_boundary"]] %||% 0L
n_multi       <- ls_counts[["multiple_zones"]]     %||% 0L
n_invalid     <- ls_counts[["invalid_geometry"]]   %||% 0L

pct_outside   <- round(100 * n_outside  / n_total, 1)
pct_straddle  <- round(100 * n_straddle / n_total, 1)
pct_multi     <- round(100 * n_multi    / n_total, 1)
pct_in_zone   <- round(100 * n_in_zone  / n_total, 1)

# Orphan zones (zones-side)
orphan_zones <- tryCatch(
  {
    bldgs_31287 <- if (sf::st_crs(buildings)$epsg == 4326L) {
      sf::st_transform(buildings, crs = 31287L)
    } else {
      buildings
    }
    zones_31287 <- if (sf::st_crs(zones)$epsg == 4326L) {
      sf::st_transform(zones, crs = 31287L)
    } else {
      zones
    }
    acd:::find_orphan_zones(zones_31287, bldgs_31287) |>
      sf::st_transform(crs = 4326L)
  },
  error = function(e) {
    # Local fallback: zones not covered by any in_zone or straddles building
    covered_zones <- joined |>
      sf::st_drop_geometry() |>
      dplyr::filter(!is.na(zone_id)) |>
      dplyr::pull(zone_id) |>
      unique()
    zones[!zones$zone_id %in% covered_zones, ]
  }
)
n_orphan <- nrow(orphan_zones)

# ── Districts loader ──────────────────────────────────────────────────────────
districts <- tryCatch({
  acd::load_acd_data("districts", source = data_source)
}, error = function(e) {
  message("load_acd_data('districts') failed; trying .cached_parquet fallback")
  districts_raw <- .cached_parquet(paste0("districts_", SNAPSHOT_DATE, ".parquet"))
  if (!is.null(districts_raw)) {
    sf::st_as_sf(districts_raw, wkt = "geometry_wkt", crs = 31287L)
  } else {
    NULL
  }
})

# Transform districts to WGS84 for leaflet (if available)
if (!is.null(districts)) {
  districts_wgs84 <- sf::st_transform(districts, crs = 4326L)
  # Identify focus districts (BEZ "20" and "21")
  districts_focus_idx <- which(
    !is.na(districts_wgs84$BEZNR) &
    as.character(districts_wgs84$BEZNR) %in% c("20", "21")
  )
} else {
  districts_wgs84 <- NULL
  districts_focus_idx <- integer(0)
}

# ── Confidence tier palette ────────────────────────────────────────────────────
tier_colors <- c(
  "A_in_zone_high"      = "#2ca02c",
  "B_in_zone_boundary"  = "#bcbd22",
  "C_outside_boundary"  = "#ff7f0e",
  "D_outside_nearby"    = "#9467bd",
  "E_outside_far"       = "#7f7f7f"
)
tier_labels <- c(
  "A_in_zone_high"      = "A. In zone (high confidence)",
  "B_in_zone_boundary"  = "B. In zone (boundary)",
  "C_outside_boundary"  = "C. Outside (boundary ≤25 m)",
  "D_outside_nearby"    = "D. Outside (nearby 25–500 m)",
  "E_outside_far"       = "E. Outside (far >500 m)"
)

# Leaflet color function for confidence_tier
pal_tier <- leaflet::colorFactor(
  palette  = unname(tier_colors),
  levels   = names(tier_colors),
  na.color = "#999999"
)

# Confidence tier summary
tier_summary <- tryCatch(
  acd::confidence_tier_summary(joined),
  error = function(e) {
    # local fallback
    if ("confidence_tier" %in% names(joined)) {
      joined |>
        sf::st_drop_geometry() |>
        dplyr::count(confidence_tier, name = "n_buildings") |>
        dplyr::mutate(
          pct   = round(100 * n_buildings / sum(n_buildings), 1),
          color = tier_colors[as.character(confidence_tier)]
        )
    } else {
      tibble::tibble(
        confidence_tier = character(0),
        n_buildings     = integer(0),
        pct             = numeric(0),
        color           = character(0)
      )
    }
  }
)

# ── Compatibility: keep .has_acd_agreement FALSE for page-6 guards ─────────────
.has_acd_agreement <- FALSE  # Stage 1a removed acd_agreement; use confidence_tier

# Placeholder for removed acd_agreement pre-computation
.ag <- function(level) {
  # acd_agreement removed in Stage 1a — returns 0 for all levels
  0L
}
n_agree_in_zone  <- 0L
n_agree_no_zone  <- 0L
n_spatial_only   <- 0L
n_acd_only       <- 0L
n_agree_total    <- 0L
n_disagree_total <- 0L

# Kept for legacy references below that will not be rendered (Page 3 tab 2 replaced)
confusion_matrix_df <- data.frame(
  `Spatial join result` = character(0), check.names = FALSE
)
acd_disagree_sf <- joined[integer(0), ]

# Placeholder for removed .ag function body (unused after Stage 1a)
.removed_ag_body <- function(level) {
  v <- integer(0)  # acd_agreement removed in Stage 1a
  if (length(v) == 0L) 0L else as.integer(v)
}

# ── Per-case subsets for page 4 ───────────────────────────────────────────────
case_in_zone <- function(df, n = 50L) {
  sub <- df[df$link_status == "in_zone", ]
  if (nrow(sub) > n) sub <- sub[sample(nrow(sub), n), ]
  sub
}

case_outside <- function(df, n = 100L) {
  sub <- df[df$link_status == "outside_any_zone", ]
  if (nrow(sub) > n) sub <- sub[sample(nrow(sub), n), ]
  sub
}

case_straddle <- function(df, n = 50L) {
  sub <- df[df$link_status == "straddles_boundary", ]
  if (nrow(sub) > n) sub <- sub[sample(nrow(sub), n), ]
  sub
}

case_multi <- function(df, n = 50L) {
  sub <- df[df$link_status == "multiple_zones", ]
  if (nrow(sub) > n) sub <- sub[sample(nrow(sub), n), ]
  sub
}

case_invalid <- function(df) {
  df[df$link_status == "invalid_geometry", ]
}

# ── Focus districts (T3) ──────────────────────────────────────────────────────
FOCUS_DISTRICTS_NAMES   <- c("Brigittenau", "Floridsdorf")
FOCUS_DISTRICTS_NUMBERS <- c(20L, 21L)
is_focus_district <- function(bez) bez %in% FOCUS_DISTRICTS_NAMES

# ── Color palette for link_status (dark-background optimised) ─────────────────
link_status_colors <- c(
  in_zone            = "#66ff66",  # bright green
  outside_any_zone   = "#cccccc",  # light grey
  straddles_boundary = "#ffaa44",  # warm orange
  multiple_zones     = "#c39bd3",  # soft purple
  orphan_zone        = "#66b3ff",  # sky blue
  invalid_geometry   = "#ff5555"   # bright red
)

link_status_labels <- c(
  in_zone            = "Clean join (in zone)",
  outside_any_zone   = "Outside any zone",
  straddles_boundary = "Straddles boundary",
  multiple_zones     = "Multiple zones",
  orphan_zone        = "Orphan zone",
  invalid_geometry   = "Invalid geometry"
)

# Leaflet color function for link_status
pal_link <- leaflet::colorFactor(
  palette = unname(link_status_colors),
  levels  = names(link_status_colors),
  na.color = "#999999"
)

# Color function for zone_type
zone_type_colors <- c(
  district_heating = "#e41a1c",
  heat_pump        = "#377eb8",
  fossil_fuel_ban  = "#984ea3"
)
pal_zone_type <- leaflet::colorFactor(
  palette  = unname(zone_type_colors),
  levels   = names(zone_type_colors),
  na.color = "#aaaaaa"
)

# Continuous palette for year_built
year_range <- range(buildings$year_built, na.rm = TRUE)
pal_year   <- leaflet::colorNumeric(
  palette = "YlOrRd",
  domain  = year_range
)

# ── Vienna map centre ─────────────────────────────────────────────────────────
vienna_center <- c(lng = 16.3725, lat = 48.2082)

# ── Story buildings / zones for page 5 ───────────────────────────────────────
# Helper: return first row where link_status matches, or NULL if none exist
.story_row <- function(df, status, fallback = NULL) {
  sub <- df[!is.na(df$link_status) & df$link_status == status, ]
  if (nrow(sub) == 0L) return(fallback)
  sub[1L, ]
}

story_clean    <- .story_row(joined, "in_zone",            joined[1L, ])
story_outside  <- .story_row(joined, "outside_any_zone",   joined[nrow(joined), ])
story_straddle <- .story_row(joined, "straddles_boundary", joined[1L, ])
story_orphan   <- if (nrow(orphan_zones) > 0L) orphan_zones[1L, ] else NULL
story_invalid  <- .story_row(joined, "invalid_geometry",   NULL)

# Compute area_ha for zones if not present
if (!"area_ha" %in% names(zones)) {
  zones_area <- zones |>
    sf::st_transform(crs = 31287L) |>
    dplyr::mutate(area_ha = round(as.numeric(sf::st_area(geometry_wkt)) / 10000, 1)) |>
    sf::st_transform(crs = 4326L)
} else {
  zones_area <- zones
}

# n_buildings per zone for the zone table
zone_bldg_counts <- joined |>
  sf::st_drop_geometry() |>
  dplyr::filter(!is.na(zone_id)) |>
  dplyr::count(zone_id, name = "n_buildings")

zone_table_df <- zones_area |>
  sf::st_drop_geometry() |>
  dplyr::left_join(zone_bldg_counts, by = "zone_id") |>
  dplyr::mutate(n_buildings = dplyr::coalesce(as.integer(n_buildings), 0L)) |>
  dplyr::select(
    zone_id, zone_type, legal_status,
    dplyr::any_of("area_ha"),
    n_buildings
  )

# Buildings per district summary for page 2
district_summary <- joined |>
  sf::st_drop_geometry() |>
  dplyr::count(district, name = "n_buildings") |>
  dplyr::left_join(
    joined |>
      sf::st_drop_geometry() |>
      dplyr::filter(!is.na(zone_id)) |>
      dplyr::count(district, name = "n_in_zone"),
    by = "district"
  ) |>
  dplyr::mutate(
    n_in_zone = dplyr::coalesce(as.integer(n_in_zone), 0L),
    pct_in_zone = round(100 * n_in_zone / n_buildings, 1)
  ) |>
  dplyr::arrange(district)

# Stats per district table (T4): focus districts first (Stage 1a: removed acd_agreement columns)
district_stats_full <- joined |>
  sf::st_drop_geometry() |>
  dplyr::count(district, name = "n_buildings") |>
  dplyr::left_join(
    joined |>
      sf::st_drop_geometry() |>
      dplyr::filter(!is.na(zone_id)) |>
      dplyr::count(district, name = "n_in_zone"),
    by = "district"
  ) |>
  dplyr::mutate(
    n_in_zone   = dplyr::coalesce(as.integer(n_in_zone), 0L),
    pct_in_zone = round(100 * n_in_zone / n_buildings, 1),
    is_focus    = district %in% FOCUS_DISTRICTS_NAMES
  ) |>
  dplyr::arrange(dplyr::desc(is_focus), dplyr::desc(n_buildings))

# decade() helper
decade <- function(year) {
  paste0(floor(year / 10) * 10, "s")
}

# ── Map sample for leaflets (cap at 5000 rows for browser performance) ─────────
# Full data is kept in `joined` for tables, summaries, and crosstabs.
# `joined_for_map` is used only for leaflet markers to avoid browser OOM.
MAP_SAMPLE_N <- 5000L
joined_for_map <- if (nrow(joined) > MAP_SAMPLE_N) {
  set.seed(42L)
  joined[sample.int(nrow(joined), MAP_SAMPLE_N), ]
} else {
  joined
}

# ── District layer helper (T4) ────────────────────────────────────────────────
# Adds district polygon outlines to a leaflet map. Focus districts (20, 21)
# use a bright yellow stroke; all others use a subtle gray stroke.
# Pass `show_layer_control = TRUE` to include a toggle in layer controls.
.add_districts_to_leaflet <- function(m, group_name = "Districts") {
  if (is.null(districts_wgs84) || nrow(districts_wgs84) == 0L) return(m)

  is_focus <- as.character(districts_wgs84$BEZNR) %in% c("20", "21")

  popup_html <- htmltools::HTML(paste0(
    "<div style='color:#fff;background:#111;padding:5px;'>",
    "<strong>Bezirk ", districts_wgs84$BEZNR, "</strong><br>",
    districts_wgs84$NAMEG,
    "</div>"
  ))

  # District outlines are decorative overlays — clicks must pass THROUGH
  # them so buildings/zones underneath retain their popups. Use label= for
  # a hover tooltip instead of a click popup.
  pass_through_opts <- leaflet::pathOptions(
    interactive = FALSE,
    pane        = "overlayPane"
  )
  highlight_off <- leaflet::highlightOptions(stroke = FALSE)

  # Non-focus districts
  m <- leaflet::addPolygons(
    m,
    data        = districts_wgs84[!is_focus, ],
    fillOpacity = 0,
    color       = "#666666",
    weight      = 1,
    opacity     = 0.7,
    group       = group_name,
    options     = pass_through_opts,
    highlight   = highlight_off,
    label       = ~ paste0("Bezirk ", BEZNR, " — ", NAMEG)
  )
  # Focus districts (20 Brigittenau, 21 Floridsdorf)
  m <- leaflet::addPolygons(
    m,
    data        = districts_wgs84[is_focus, ],
    fillOpacity = 0,
    color       = "#ffff00",
    weight      = 4,
    opacity     = 0.9,
    group       = group_name,
    options     = pass_through_opts,
    highlight   = highlight_off,
    label       = ~ paste0("Bezirk ", BEZNR, " (focus) — ", NAMEG)
  )
  m
}
```

```{=html}
<script>
document.addEventListener('DOMContentLoaded', function() {
  document.querySelectorAll('[data-bs-toggle="tab"]').forEach(function(tab) {
    tab.addEventListener('shown.bs.tab', function() {
      window.dispatchEvent(new Event('resize'));
    });
  });
  setTimeout(function() { window.dispatchEvent(new Event('resize')); }, 500);
  setTimeout(function() { window.dispatchEvent(new Event('resize')); }, 2000);
});
</script>
```

# Source A — Zones

## Row

::: {.card title="Zone Map"}

```{r}
#| label: fig-zone-map

zones_wgs84 <- sf::st_transform(zones, crs = 4326L)

zone_pal <- leaflet::colorFactor(
  palette  = c("#e41a1c", "#377eb8", "#984ea3", "#ff7f0e"),
  domain   = zones_wgs84$zone_type,
  na.color = "#aaaaaa"
)

# Focus-district mask for zone outlines (T3)
focus_zone_ids <- if ("district" %in% names(zones_wgs84)) {
  zones_wgs84$zone_id[zones_wgs84$district %in% FOCUS_DISTRICTS_NAMES]
} else {
  character(0)
}

m_zone <- leaflet(data = zones_wgs84, height = "500px") |>
  leaflet::addProviderTiles(
    leaflet::providers$CartoDB.DarkMatter,
    options = leaflet::providerTileOptions(opacity = 0.95)
  ) |>
  leaflet::addPolygons(
    fillColor   = "#999999",
    fillOpacity = 0.30,
    color       = "#CC0000",
    weight      = 2,
    opacity     = 1,
    group       = "Zones",
    popup       = ~ htmltools::HTML(paste0(
      "<div style='color:#fff;background:#000;padding:6px;'>",
      "<strong>Zone: </strong>", zone_id, "<br>",
      "<strong>Type: </strong>", zone_type, "<br>",
      "<strong>Legal status: </strong>", legal_status,
      "</div>"
    )),
    label = ~ paste0(zone_id, " (", zone_type, ")")
  ) |>
  leaflet::addLegend(
    position = "bottomright",
    pal      = zone_pal,
    values   = ~ zone_type,
    title    = "Zone type",
    opacity  = 0.9
  ) |>
  leaflet::setView(
    lng  = vienna_center[["lng"]],
    lat  = vienna_center[["lat"]],
    zoom = 12L
  )

m_zone <- .add_districts_to_leaflet(m_zone)
m_zone <- leaflet::addLayersControl(
  m_zone,
  overlayGroups = c("Zones", "Districts"),
  options       = leaflet::layersControlOptions(collapsed = TRUE)
)
m_zone
```

:::

## Row

::: {.card title="Zone Attributes"}

```{r}
#| label: tbl-zone-attributes

DT::datatable(
  zone_table_df,
  rownames = FALSE,
  filter   = "top",
  colnames = c("Zone ID", "Type", "Legal status", "Area (ha)", "N buildings"),
  options  = list(
    pageLength = 15,
    scrollX    = TRUE,
    order      = list(list(4L, "desc"))
  )
)
```

:::

::: {.card title="Notes"}

### ACD Zones — Klimaschutzgebiete

Vienna's *Energieraumplanung* (Energy Area Planning, ACD) designates geographic
zones where specific heating and energy-system rules apply to buildings.

**Layer:** `ogdwien:ENERGIEPLANERPOGD`
**Provider:** [Vienna Open Government Data WFS](https://data.wien.gv.at/daten/geo)
**Snapshot date:** `r SNAPSHOT_DATE`
**Total zones in this snapshot:** `r n_zones`

Zone types:

| Type | Legal rule |
|------|-----------|
| `district_heating` | Buildings must connect to district heating network |
| `heat_pump` | Heat-pump installations required; fossil fuel banned |
| `fossil_fuel_ban` | Existing fossil-fuel heating must be retired |

Source columns used from the WFS layer: `ERPLABEL` (zone ID),
`ERPTYP` (zone type), `ERPRECHTSTAT` (legal status).

:::

# Source B — Buildings

## Row

::: {.card title="Building Sample (first 200)"}

```{r}
#| label: fig-building-sample

bldg_sample <- if (nrow(buildings) > 200L) buildings[seq_len(200L), ] else buildings

m_bldg <- leaflet(data = bldg_sample, height = "500px") |>
  leaflet::addProviderTiles(
    leaflet::providers$CartoDB.DarkMatter,
    options = leaflet::providerTileOptions(opacity = 0.95)
  ) |>
  leaflet::addCircleMarkers(
    radius      = 4,
    fillColor   = ~ pal_year(year_built),
    fillOpacity = 0.60,
    color       = "#999999",
    weight      = 0.5,
    opacity     = 0.7,
    group       = "Buildings",
    popup       = ~ htmltools::HTML(paste0(
      "<div style='color:#fff;background:#000;padding:6px;'>",
      "<strong>Address: </strong>", address, "<br>",
      "<strong>Year built: </strong>", year_built, "<br>",
      "<strong>Height: </strong>", height_m, " m",
      "</div>"
    ))
  ) |>
  leaflet::addLegend(
    position = "bottomright",
    pal      = pal_year,
    values   = ~ year_built,
    title    = "Year built",
    opacity  = 0.9
  ) |>
  leaflet::setView(
    lng  = vienna_center[["lng"]],
    lat  = vienna_center[["lat"]],
    zoom = 12L
  )
m_bldg <- .add_districts_to_leaflet(m_bldg)
m_bldg <- leaflet::addLayersControl(
  m_bldg,
  overlayGroups = c("Buildings", "Districts"),
  options       = leaflet::layersControlOptions(collapsed = TRUE)
)
m_bldg
```

:::

## Row

::: {.card title="Buildings by District"}

```{r}
#| label: tbl-district-summary

district_summary |>
  gt::gt() |>
  gt::cols_label(
    district    = "District",
    n_buildings = "Total",
    n_in_zone   = "In a zone",
    pct_in_zone = "% in zone"
  ) |>
  gt::fmt_number(columns = c(n_buildings, n_in_zone), decimals = 0) |>
  gt::fmt_number(columns = pct_in_zone, decimals = 1) |>
  gt::data_color(
    columns = pct_in_zone,
    palette = c("#fee0d2", "#de2d26")
  ) |>
  gt::tab_caption(
    caption = paste0(
      "Buildings per district as of snapshot ", SNAPSHOT_DATE,
      ". 'In a zone' counts buildings with link_status != outside_any_zone."
    )
  )
```

:::

::: {.card title="Year-Built Distribution"}

```{r}
#| label: fig-year-hist

yb <- buildings$year_built[!is.na(buildings$year_built)]
hist(
  yb,
  breaks = 30,
  col    = "#377eb8",
  border = "#1a5276",
  xlab   = "Year built",
  ylab   = "Number of buildings",
  main   = paste0(
    "Year-Built Distribution (N = ", length(yb), ", snapshot ", SNAPSHOT_DATE, ")"
  ),
  las    = 1
)
```

:::

::: {.card title="Notes"}

### Vienna Building Inventory

**Layer:** `ogdwien:GEBAEUDEOGD`
**Provider:** [Vienna Open Government Data WFS](https://data.wien.gv.at/daten/geo)
**Snapshot date:** `r SNAPSHOT_DATE`
**Buildings in this snapshot:** `r format(n_total, big.mark = ",")` (full WFS layer, `max_features = NULL`)

For rendering speed the map above shows the first
`r min(nrow(buildings), 200L)` buildings only.

Column mapping from WFS: `OBJECTID` → `building_id`, `BAUJAHR` → `year_built`,
`HOEHE` / `GEBHOEHE` → `height_m`, `ADRESSE` → `address`.

:::

::: {.card title="Stats per District (T4)"}

```{r}
#| label: tbl-district-stats-full

district_stats_full |>
  dplyr::select(-is_focus) |>
  gt::gt() |>
  gt::cols_label(
    district    = "District",
    n_buildings = "N buildings",
    n_in_zone   = "N in zone",
    pct_in_zone = "% in zone"
  ) |>
  gt::fmt_number(columns = c(n_buildings, n_in_zone), decimals = 0) |>
  gt::fmt_number(columns = pct_in_zone, decimals = 1) |>
  gt::tab_style(
    style     = gt::cell_fill(color = "#fff8dc"),
    locations = gt::cells_body(
      rows = district %in% FOCUS_DISTRICTS_NAMES
    )
  ) |>
  gt::data_color(
    columns = pct_in_zone,
    palette = c("#fee0d2", "#de2d26")
  ) |>
  gt::tab_caption(
    caption = paste0(
      "Buildings per district (snapshot: ", SNAPSHOT_DATE, "). ",
      "Focus districts Brigittenau (20) and Floridsdorf (21) are highlighted and sorted first. ",
      "N in zone = buildings whose centroid falls within at least one ACD energy-planning zone ",
      "(link_status != outside_any_zone). ",
      "Source: ogdwien:GEBAEUDEINFOOGD + ogdwien:ENERGIEPLANERPOGD. ",
      "N = ", format(n_total, big.mark = ","), " buildings across ",
      nrow(district_stats_full), " districts."
    )
  )
```

:::

## Row

::: {.card title="OSM vs OGD Coverage (Stage 2 — issue #14)"}

```{r}
#| label: tbl-osm-coverage

# Load Stage 2 comparison CSV (written by analysis/stage_2/merge_comparison.R)
.osm_cov_path <- system.file(
  "extdata", "osm_coverage_by_district.csv", package = "acd"
)
# Fall back to analysis output if not yet bundled in inst/extdata
if (!nzchar(.osm_cov_path) || !file.exists(.osm_cov_path)) {
  .osm_cov_path <- file.path(
    tryCatch(find.package("acd"), error = function(e) "."),
    "inst", "extdata", "osm_coverage_by_district.csv"
  )
}
if (!file.exists(.osm_cov_path)) {
  .osm_cov_path <- "analysis/stage_2/coverage_by_district.csv"
}

if (file.exists(.osm_cov_path)) {
  osm_cov <- read.csv(.osm_cov_path, stringsAsFactors = FALSE)

  osm_total_full <- tryCatch({
    .smry_path <- sub("coverage_by_district", "osm_total_summary", .osm_cov_path)
    if (!file.exists(.smry_path)) {
      .smry_path <- "analysis/stage_2/osm_total_summary.csv"
    }
    read.csv(.smry_path)$n_total[1]
  }, error = function(e) NA_integer_)

  ogd_grand_total <- sum(osm_cov$n_ogd, na.rm = TRUE)
  osm_in_dist     <- sum(osm_cov$n_osm, na.rm = TRUE)

  # Show focus districts first, then top-10 by ratio
  top_rows <- osm_cov |>
    dplyr::arrange(dplyr::desc(is_focus), dplyr::desc(ratio_osm_to_ogd)) |>
    dplyr::slice_head(n = 13)

  focus_row_idx <- which(top_rows$is_focus)

  top_rows |>
    dplyr::mutate(is_focus = ifelse(is_focus, "★", "")) |>
    dplyr::select(district, district_name, n_ogd, n_osm, ratio_osm_to_ogd, is_focus) |>
    gt::gt() |>
    gt::cols_label(
      district         = "Bez.",
      district_name    = "District",
      n_ogd            = "OGD (GEBAEUDEINFOOGD)",
      n_osm            = "OSM bbox count",
      ratio_osm_to_ogd = "OSM / OGD",
      is_focus         = "Focus"
    ) |>
    gt::fmt_number(columns = c(n_ogd, n_osm), decimals = 0) |>
    gt::fmt_number(columns = ratio_osm_to_ogd, decimals = 1) |>
    gt::tab_style(
      style     = gt::cell_fill(color = "#fff8dc"),
      locations = gt::cells_body(rows = focus_row_idx)
    ) |>
    gt::data_color(
      columns = ratio_osm_to_ogd,
      palette = c("#deebf7", "#08519c")
    ) |>
    gt::tab_footnote(
      footnote = paste0(
        "OSM full-city total (bbox query, 2026-05-01): ",
        format(osm_total_full, big.mark = ","), " buildings. ",
        "OGD city-wide total: ",
        format(ogd_grand_total, big.mark = ","), ". ",
        "Overall OSM/OGD ratio: ",
        round(osm_total_full / max(ogd_grand_total, 1), 1), "x. ",
        "OGD covers ~",
        round(100 * ogd_grand_total / max(osm_total_full, 1), 0),
        "% of the OSM building count. ",
        "OSM per-district counts use bounding boxes (not polygon-clipped), ",
        "so outer-district ratios are inflated. ",
        "Source: OpenStreetMap contributors, ODbL 1.0."
      )
    ) |>
    gt::tab_caption(
      caption = paste0(
        "OGD vs OSM building counts per district (Stage 2, issue #14). ",
        "OSM has ", round(osm_total_full / max(ogd_grand_total, 1), 1), "x more ",
        "buildings city-wide than GEBAEUDEINFOOGD. ",
        "Focus districts 20 (Brigittenau) and 21 (Floridsdorf) highlighted. ",
        "Top 13 districts by ratio shown."
      )
    )
} else {
  gt::gt(data.frame(Note = paste0(
    "Stage 2 coverage data not yet available. ",
    "Run analysis/stage_2/merge_comparison.R to generate coverage_by_district.csv."
  )))
}
```

:::

# Joined: Buildings × Zones

## Row

::: {.panel-tabset}

### Joined map

::: {.card title="Joined Map"}

```{r}
#| label: fig-joined-map

zones_wgs84_join <- sf::st_transform(zones, crs = 4326L)

# T6: Investigate ~100 points — joined_for_map is 5000 rows but uses
# clusterOptions to handle density; actual sample size is MAP_SAMPLE_N.
# Use markerClusterOptions to show all points without blobs.
leaflet(height = 500) |>
  leaflet::addProviderTiles(
    leaflet::providers$CartoDB.DarkMatter,
    options = leaflet::providerTileOptions(opacity = 0.95)
  ) |>
  leaflet::addPolygons(
    data        = zones_wgs84_join,
    fillColor   = "#999999",
    fillOpacity = 0.30,
    color       = "#CC0000",
    weight      = 2,
    opacity     = 0.8,
    group       = "Zones",
    popup       = ~ htmltools::HTML(paste0(
      "<div style='color:#fff;background:#000;padding:6px;'>",
      "<strong>Zone: </strong>", zone_id,
      "</div>"
    ))
  ) |>
  leaflet::addMarkers(
    data  = joined_for_map,
    group = "Buildings",
    clusterOptions = leaflet::markerClusterOptions(),
    popup = ~ htmltools::HTML(paste0(
      "<div style='color:#fff;background:#000;padding:6px;'>",
      "<strong>Building: </strong>", building_id, "<br>",
      "<strong>Status: </strong>", link_status, "<br>",
      "<strong>Zone: </strong>", dplyr::coalesce(zone_id, "none"),
      "</div>"
    ))
  ) |>
  leaflet::addLegend(
    position = "bottomright",
    pal      = pal_link,
    values   = joined$link_status,
    title    = "Link status",
    opacity  = 0.9
  ) |>
  leaflet::addLayersControl(
    overlayGroups = c("Buildings", "Zones", "Districts"),
    options       = leaflet::layersControlOptions(collapsed = FALSE)
  ) |>
  leaflet::setView(
    lng  = vienna_center[["lng"]],
    lat  = vienna_center[["lat"]],
    zoom = 12L
  ) |>
  (\(m) .add_districts_to_leaflet(m))()
```

**Map caption:** Spatial overlay of `r format(nrow(joined_for_map), big.mark = ",")` sampled buildings (coloured by `link_status`) on `r n_zones` ACD climate-zone polygons (red outlines). Variables: `link_status` (building–zone linkage outcome, 6 levels), `zone_id` (ACD polygon identifier). Colour: bright green = in-zone; light grey = outside all zones; warm orange = straddles boundary; soft purple = multiple overlapping zones; sky blue = orphan zone; bright red = invalid geometry. Yellow outlines mark focus districts Brigittenau (20) and Floridsdorf (21). `r round(100 * n_in_zone / n_total, 1)`% of the `r format(n_total, big.mark = ",")` buildings fall within at least one ACD zone. Source: ogdwien:GEBAEUDEINFOOGD + ogdwien:ENERGIEPLANERPOGD, snapshot `r SNAPSHOT_DATE`. See tbl-link-summary for full count breakdown; see Confidence tab for tier analysis.

:::

### Confidence: how sure are we?

::: {.card title="Tier distribution"}

```{r}
#| label: tbl-p3-tier-dist

if (nrow(tier_summary) > 0L) {
  tier_display <- tier_summary |>
    dplyr::mutate(
      tier_label = tier_labels[as.character(confidence_tier)],
      tier_label = dplyr::coalesce(tier_label, as.character(confidence_tier)),
      # Blank cell text — tab_style fills the cell BACKGROUND so the text
      # would just be a redundant hex string layered on the same colour.
      color      = ""
    ) |>
    dplyr::select(tier_label, n_buildings, pct, color)

  # Pre-compute row indices for each tier (for gt row styling)
  tier_codes  <- as.character(tier_summary$confidence_tier)
  row_A <- which(grepl("^A", tier_codes))
  row_B <- which(grepl("^B", tier_codes))
  row_C <- which(grepl("^C", tier_codes))
  row_D <- which(grepl("^D", tier_codes))
  row_E <- which(grepl("^E", tier_codes))

  tbl_tier_summary <- tier_display |>
    gt::gt() |>
    gt::cols_label(
      tier_label  = "Confidence tier",
      n_buildings = "N buildings",
      pct         = "% of total",
      color       = "Color"
    ) |>
    gt::fmt_number(columns = n_buildings, decimals = 0) |>
    gt::fmt_number(columns = pct, decimals = 1) |>
    gt::cols_width(color ~ gt::px(60))

  if (length(row_A)) tbl_tier_summary <- gt::tab_style(
    tbl_tier_summary,
    style = gt::cell_fill(color = "#2ca02c"),
    locations = gt::cells_body(columns = color, rows = row_A)
  )
  if (length(row_B)) tbl_tier_summary <- gt::tab_style(
    tbl_tier_summary,
    style = gt::cell_fill(color = "#bcbd22"),
    locations = gt::cells_body(columns = color, rows = row_B)
  )
  if (length(row_C)) tbl_tier_summary <- gt::tab_style(
    tbl_tier_summary,
    style = gt::cell_fill(color = "#ff7f0e"),
    locations = gt::cells_body(columns = color, rows = row_C)
  )
  if (length(row_D)) tbl_tier_summary <- gt::tab_style(
    tbl_tier_summary,
    style = gt::cell_fill(color = "#9467bd"),
    locations = gt::cells_body(columns = color, rows = row_D)
  )
  if (length(row_E)) tbl_tier_summary <- gt::tab_style(
    tbl_tier_summary,
    style = gt::cell_fill(color = "#7f7f7f"),
    locations = gt::cells_body(columns = color, rows = row_E)
  )

  tbl_tier_summary |>
    gt::tab_caption(
      caption = paste0(
        "Confidence tier distribution for ", format(n_total, big.mark = ","),
        " Vienna buildings, snapshot ", SNAPSHOT_DATE, ". ",
        "The 5 tiers are based on centroid-to-zone-boundary distance ",
        "(EPSG:31287 metres), NOT the original broken 'city ACD vs spatial' comparison ",
        "which Stage 0 falsified. ",
        "Tiers A and B (inside any zone): ",
        sum(tier_summary$n_buildings[startsWith(as.character(tier_summary$confidence_tier), "A") |
                                     startsWith(as.character(tier_summary$confidence_tier), "B")]),
        " buildings. ",
        "Tiers C–E (outside any zone): ",
        sum(tier_summary$n_buildings[startsWith(as.character(tier_summary$confidence_tier), "C") |
                                     startsWith(as.character(tier_summary$confidence_tier), "D") |
                                     startsWith(as.character(tier_summary$confidence_tier), "E")]),
        " buildings. ",
        "Source: join_buildings_zones() using sf::st_distance() on EPSG:31287 coordinates. ",
        "See analysis/stage_0/REPORT.md for the Stage 0 falsification finding."
      )
    )
} else {
  cat("confidence_tier column not available — tier distribution unavailable for this snapshot.")
}
```

:::

::: {.card title="Map of confidence tiers"}

```{r}
#| label: fig-p3-tier-map

if ("confidence_tier" %in% names(joined_for_map)) {
  m_tier <- leaflet(height = 500) |>
    leaflet::addProviderTiles(
      leaflet::providers$CartoDB.DarkMatter,
      options = leaflet::providerTileOptions(opacity = 0.95)
    ) |>
    leaflet::addPolygons(
      data        = sf::st_transform(zones, crs = 4326L),
      fillColor   = "#555555",
      fillOpacity = 0.20,
      color       = "#CC0000",
      weight      = 1.5,
      opacity     = 0.7,
      group       = "Zones",
      popup       = ~ htmltools::HTML(paste0(
        "<div style='color:#fff;background:#000;padding:5px;'>",
        "<strong>Zone: </strong>", zone_id,
        "</div>"
      ))
    ) |>
    leaflet::addCircleMarkers(
      data        = joined_for_map,
      radius      = 4,
      fillColor   = ~ pal_tier(as.character(confidence_tier)),
      fillOpacity = 0.80,
      color       = "#333333",
      weight      = 0.4,
      opacity     = 0.6,
      group       = "Buildings",
      popup       = ~ htmltools::HTML(paste0(
        "<div style='color:#fff;background:#000;padding:5px;'>",
        "<strong>Building: </strong>", building_id, "<br>",
        "<strong>Tier: </strong>", confidence_tier, "<br>",
        "<strong>Distance: </strong>",
        if ("nearest_zone_dist_m" %in% names(joined_for_map))
          paste0(round(nearest_zone_dist_m, 0), " m")
        else "(unavailable)",
        "</div>"
      ))
    ) |>
    leaflet::addLegend(
      position = "bottomright",
      pal      = pal_tier,
      values   = names(tier_colors),
      title    = "Confidence tier",
      opacity  = 0.9
    ) |>
    leaflet::setView(
      lng  = vienna_center[["lng"]],
      lat  = vienna_center[["lat"]],
      zoom = 12L
    )
  m_tier <- .add_districts_to_leaflet(m_tier)
  m_tier <- leaflet::addLayersControl(
    m_tier,
    overlayGroups = c("Buildings", "Zones", "Districts"),
    options       = leaflet::layersControlOptions(collapsed = FALSE)
  )
  m_tier
} else {
  cat("confidence_tier column not present — map unavailable for this snapshot.")
}
```

:::

::: {.card title="Tier explanation — Stage 0 finding"}

### What the 5 tiers mean

The `confidence_tier` column replaces the defunct `acd_agreement` metric.
Stage 0 (falsification study, 2026-05-01) found that the original cross-validation
was based on a **category error**: the `ACD` field in `GEBAEUDEINFOOGD` is a
building registry number (Adress-Code), not a zone-membership indicator. Every
building has this field populated, making the old "city says yes" comparison
meaningless.

See the full finding in
[analysis/stage_0/REPORT.md](https://github.com/JohnGavin/acd_area_climate_design/blob/main/analysis/stage_0/REPORT.md).

The **replacement metric** is centroid-to-boundary distance:

| Tier | Definition | Threshold |
|------|-----------|-----------|
| A — in zone (high) | Inside zone, centroid > 5 m from boundary | >5 m inside |
| B — in zone (boundary) | Inside zone, centroid ≤ 5 m from boundary | ≤5 m inside |
| C — outside (boundary) | Outside, within 25 m of nearest zone | ≤25 m outside |
| D — outside (nearby) | Outside, 25–500 m from nearest zone | 25–500 m outside |
| E — outside (far) | Outside, > 500 m from any zone | >500 m outside |

Tiers A+B = `r sum(tier_summary$n_buildings[startsWith(as.character(tier_summary$confidence_tier), "A") | startsWith(as.character(tier_summary$confidence_tier), "B")])` buildings are confidently inside a zone.
Tiers C–E = `r sum(tier_summary$n_buildings[startsWith(as.character(tier_summary$confidence_tier), "C") | startsWith(as.character(tier_summary$confidence_tier), "D") | startsWith(as.character(tier_summary$confidence_tier), "E")])` buildings are outside zones (varying proximity).

:::

:::

## Row

::: {.card title="Buildings by Decade"}

```{r}
#| label: tbl-decade-marginals

decade_marginals <- joined |>
  sf::st_drop_geometry() |>
  dplyr::filter(!is.na(year_built)) |>
  dplyr::mutate(decade = decade(year_built)) |>
  dplyr::count(decade, name = "total_buildings") |>
  dplyr::left_join(
    joined |>
      sf::st_drop_geometry() |>
      dplyr::filter(!is.na(year_built), !is.na(zone_id)) |>
      dplyr::mutate(decade = decade(year_built)) |>
      dplyr::count(decade, name = "n_in_zone"),
    by = "decade"
  ) |>
  dplyr::mutate(
    n_in_zone   = dplyr::coalesce(as.integer(n_in_zone), 0L),
    pct_in_zone = round(100 * n_in_zone / total_buildings, 1)
  ) |>
  dplyr::arrange(dplyr::desc(decade)) |>
  dplyr::filter(total_buildings > 0L)

decade_marginals |>
  gt::gt() |>
  gt::cols_label(
    decade          = "Decade",
    total_buildings = "Total buildings",
    n_in_zone       = "In a zone",
    pct_in_zone     = "% in zone"
  ) |>
  gt::fmt_number(columns = c(total_buildings, n_in_zone), decimals = 0) |>
  gt::fmt_number(columns = pct_in_zone, decimals = 1) |>
  gt::data_color(
    columns = pct_in_zone,
    palette = c("#fee0d2", "#de2d26")
  ) |>
  gt::tab_caption(
    caption = paste0(
      "Buildings per construction decade and zone coverage. ",
      "N = ", n_total, " buildings, snapshot ", SNAPSHOT_DATE, ". ",
      "Pct in zone = buildings with any zone assignment (link_status != outside_any_zone)."
    )
  )
```

:::

::: {.card title="Link Status Distribution"}

```{r}
#| label: tbl-link-summary

link_summary |>
  dplyr::mutate(
    color_hex = link_status_colors[as.character(link_status)],
    label     = link_status_labels[as.character(link_status)]
  ) |>
  dplyr::select(label, n_buildings, pct) |>
  gt::gt() |>
  gt::cols_label(
    label       = "Link status",
    n_buildings = "N buildings",
    pct         = "% of total"
  ) |>
  gt::fmt_number(columns = n_buildings, decimals = 0) |>
  gt::fmt_number(columns = pct, decimals = 1) |>
  gt::data_color(
    columns = n_buildings,
    palette = c("#e5f5e0", "#31a354")
  ) |>
  gt::tab_caption(
    caption = paste0(
      "Distribution of building-zone linkage outcomes. ",
      "Snapshot: ", SNAPSHOT_DATE, ". Total buildings: ", n_total, "."
    )
  )
```

:::

# Missing links — step by step

## Row

::: {.panel-tabset}

### How this works

**What this page does:** Each tab below shows buildings in one specific linkage state — from the cleanest outcome (a building fully inside a single zone) to the most problematic (invalid geometry). Each tab has a map and an explanation of what the state means.

**Tabs on this page:**

- **Clean join (in_zone):** Buildings whose footprint is entirely within a single ACD zone — the happy path.
- **Outside any zone (legitimate):** Buildings with no ACD overlay — not errors; simply outside Vienna's energy-planning area.
- **Straddles boundary:** Buildings whose footprint crosses a zone edge. Assignment is ambiguous; centroid-based join assigns to one zone.
- **Multiple overlapping zones:** Buildings whose centroid falls inside two or more overlapping zone polygons simultaneously.
- **Orphan zones (no buildings):** ACD zone polygons containing zero building centroids — possible data lag or digitisation error.
- **Invalid geometry:** Buildings with self-intersecting footprints repaired by `sf::st_make_valid()`.
- **CRS mismatch:** Why coordinate reference systems matter and what the pipeline does to prevent silent mismatches.
- **Snapshot lag:** How the pinned snapshot date governs temporal coverage.

**How the pipeline works:**

```
Buildings (60k) → centroid → spatial intersect → Zones (416)
→ primary zone assignment → link_status enum
```

Note: we are not computing a join here; we are **diagnosing a join** — each tab surfaces a class of outcomes that the join produces, so analysts can decide how to handle each case in production.

### Clean join (in_zone)

**What this is:** Buildings whose footprint is fully contained in a single ACD zone polygon — the canonical happy-path outcome.
**Why it matters:** These buildings have an unambiguous legal assignment; the downstream rule lookup is a simple table join on `zone_id` with no human review required.
**What to do with this:** No action needed. Use this set as a positive control when validating your join pipeline against future snapshots.

```{r}
#| label: fig-case-in-zone

sub <- case_in_zone(joined, n = 50L)

leaflet(data = sub, height = "420px") |>
  leaflet::addProviderTiles(
    leaflet::providers$CartoDB.DarkMatter,
    options = leaflet::providerTileOptions(opacity = 0.95)
  ) |>
  leaflet::addCircleMarkers(
    radius      = 5,
    fillColor   = link_status_colors[["in_zone"]],
    fillOpacity = 0.85,
    color       = "#1a5276",
    weight      = 0.8,
    popup       = ~ paste0(
      "<strong>Building: </strong>", building_id, "<br>",
      "<strong>Zone: </strong>", zone_id, "<br>",
      "<strong>Overlap fraction: </strong>",
      round(overlap_fraction, 3)
    )
  ) |>
  leaflet::setView(
    lng  = vienna_center[["lng"]],
    lat  = vienna_center[["lat"]],
    zoom = 12L
  )
```

**<span class='good'>HEALTHY</span> — Clean join: building footprint fully
contained in a single zone polygon.**

The building centroid falls inside exactly one zone, and the overlap fraction
is >= 0.95 (at least 95% of the footprint is covered). This is the most
common, unambiguous case — the link is legally unequivocal and the downstream
rule lookup is trivial.

N (this snapshot): **`r n_in_zone`** buildings (`r pct_in_zone`% of total).

### Outside any zone (legitimate)

**What this is:** Buildings with no ACD zone overlay — the building record is valid and geocoded, but no energy-planning zone covers this address.
**Why it matters:** These are not errors; they are buildings to which Vienna's ACD rules do not currently apply. Conflating them with join failures would inflate error counts.
**What to do with this:** Exclude from ACD compliance reporting. Re-check if a zone is later drawn over this area — a snapshot refresh would reclassify these buildings.

```{r}
#| label: fig-case-outside

sub_out <- case_outside(joined, n = 100L)

leaflet(data = sub_out, height = "420px") |>
  leaflet::addProviderTiles(
    leaflet::providers$CartoDB.DarkMatter,
    options = leaflet::providerTileOptions(opacity = 0.95)
  ) |>
  leaflet::addCircleMarkers(
    radius      = 4,
    fillColor   = link_status_colors[["outside_any_zone"]],
    fillOpacity = 0.80,
    color       = "#555555",
    weight      = 0.5,
    popup       = ~ paste0(
      "<strong>Building: </strong>", building_id, "<br>",
      "<strong>District: </strong>", dplyr::coalesce(district, "—"), "<br>",
      "<strong>Year built: </strong>", year_built
    )
  ) |>
  leaflet::setView(
    lng  = vienna_center[["lng"]],
    lat  = vienna_center[["lat"]],
    zoom = 11L
  )
```

**<span class='warn'>INFORMATIONAL</span> — Building outside any ACD zone.**

The building footprint is valid, geocoding succeeded, and the building is in
the official inventory — it simply has no ACD zone overlay at this location.
**This is information, not corruption.** About `r pct_outside`% of Vienna
buildings fall here.

Distinguish from missing data: the building record is complete; the absence
of a zone merely means no ACD energy-planning rule applies to this address.
Common on Vienna's outer districts and undeveloped fringes.

N (this snapshot): **`r n_outside`** buildings (`r pct_outside`% of total).

### Straddles boundary

**What this is:** Buildings whose footprint overlaps a zone boundary — the building is partly inside and partly outside a zone, with an overlap fraction strictly between 5% and 95%.
**Why it matters:** The current centroid-based join assigns only one zone, potentially misrepresenting legal obligations for the portion of the building in a different zone.
**What to do with this:** Escalate to manual review in production. Replace single `zone_id` with a list-column of all overlapping zones plus their respective overlap fractions.

```{r}
#| label: fig-case-straddle
#| results: asis

sub_str <- case_straddle(joined, n = 50L)

if (nrow(sub_str) == 0L) {
  cat("> **No buildings in this category in the current snapshot.** The straddling-boundary check returns 0 hits on the full Vienna data — the centroid-based join produces unambiguous assignments at this resolution. If the buffer threshold widens (e.g. 25 m), some buildings would migrate from `outside_any_zone` to a near-boundary cohort; see the **Confidence** tab for the distance-based tier breakdown.\n")
} else {
  print(leaflet(data = sub_str, height = "420px") |>
    leaflet::addProviderTiles(
      leaflet::providers$CartoDB.DarkMatter,
      options = leaflet::providerTileOptions(opacity = 0.95)
    ) |>
    leaflet::addCircleMarkers(
      radius      = 5,
      fillColor   = link_status_colors[["straddles_boundary"]],
      fillOpacity = 0.85,
      color       = "#8B4513",
      weight      = 0.8,
      popup       = ~ paste0(
        "<strong>Building: </strong>", building_id, "<br>",
        "<strong>Zone: </strong>", dplyr::coalesce(zone_id, "—"), "<br>",
        "<strong>Overlap fraction: </strong>",
        round(overlap_fraction, 3)
      )
    ) |>
    leaflet::setView(
      lng  = vienna_center[["lng"]],
      lat  = vienna_center[["lat"]],
      zoom = 12L
    ))
}
```

**<span class='warn'>AMBIGUOUS</span> — Building footprint spans two or more
zones (overlap fraction between 5% and 95%).**

The POC uses centroid-based assignment: the building is attributed to whichever
zone contains its centroid. However, legal applicability may cover both zones
for different parts of the building.

**Production fix recommended:** Replace the single `zone_id` with a
list-column of all overlapping zones plus their respective overlap fractions,
and escalate ambiguous cases (overlap fraction 20%–80%) to manual review.

Sample size (this snapshot): **`r n_straddle`** buildings
(`r pct_straddle`% of total).

### Multiple overlapping zones

**What this is:** Buildings whose centroid falls simultaneously inside two or more zone polygons — either legitimate layered zoning or a digitisation error where mutually exclusive polygons overlap.
**Why it matters:** The POC reports only the first matched zone; multiple applicable rules may be silently dropped, creating compliance gaps.
**What to do with this:** Audit each case. If zones are intentionally layered (e.g., district-heating zone AND fossil-fuel ban overlap), store both rules. If the overlap is an error, report to the data provider.

```{r}
#| label: fig-case-multi
#| results: asis

sub_mul <- case_multi(joined, n = 50L)

if (nrow(sub_mul) == 0L) {
  cat("> **No buildings in this category in the current snapshot.** The full Vienna data has 0 buildings whose centroid falls inside two or more zones simultaneously. Layered zoning would produce overlapping polygons; the current Energieraumplan publication does not. If a future snapshot introduces overlap, this tab will populate.\n")
} else {
  print(leaflet(data = sub_mul, height = "420px") |>
    leaflet::addProviderTiles(
      leaflet::providers$CartoDB.DarkMatter,
      options = leaflet::providerTileOptions(opacity = 0.95)
    ) |>
    leaflet::addCircleMarkers(
      radius      = 5,
      fillColor   = link_status_colors[["multiple_zones"]],
      fillOpacity = 0.85,
      color       = "#4B0082",
      weight      = 0.8,
      popup       = ~ paste0(
        "<strong>Building: </strong>", building_id, "<br>",
        "<strong>Primary zone: </strong>",
        dplyr::coalesce(zone_id, "—"), "<br>",
        "<strong>N overlapping zones: </strong>", n_overlapping_zones
      )
    ) |>
    leaflet::setView(
      lng  = vienna_center[["lng"]],
      lat  = vienna_center[["lat"]],
      zoom = 12L
    ))
}
```

**<span class='warn'>INSPECT</span> — Building centroid falls inside two or
more zone polygons simultaneously.**

Two possible explanations:

1. **Legitimate layering**: Vienna may designate the same area under multiple
   rule frameworks simultaneously (e.g., district-heating zone AND a
   fossil-fuel ban zone covering the same geometry). In this case both rules
   apply and must be stored.

2. **Data error**: Zone polygons that should be mutually exclusive may have
   overlapping geometries due to digitisation errors. Inspect each case.

The POC reports only the first matched zone. Production must store all
overlapping zones as a list-column.

N (this snapshot): **`r n_multi`** buildings (`r pct_multi`% of total).

### Orphan zones (no buildings)

**What this is:** ACD zone polygons that contain zero building centroids — the zone exists in the energy plan but no building in the current snapshot falls inside it.
**Why it matters:** An orphan zone may be (a) newly designated over undeveloped land, (b) an inventory lag where buildings exist but are not yet catalogued, or (c) a polygon digitisation error.
**What to do with this:** Cross-reference each orphan against current satellite imagery before concluding it is a data error. Do not conflate with `outside_any_zone` buildings — the zone is real; the building list may be incomplete.

```{r}
#| label: fig-case-orphan

if (nrow(orphan_zones) > 0L) {
  leaflet(data = sf::st_transform(orphan_zones, crs = 4326L), height = "420px") |>
    leaflet::addProviderTiles(
    leaflet::providers$CartoDB.DarkMatter,
    options = leaflet::providerTileOptions(opacity = 0.95)
  ) |>
    leaflet::addPolygons(
      fillColor   = link_status_colors[["orphan_zone"]],
      fillOpacity = 0.35,
      color       = link_status_colors[["orphan_zone"]],
      weight      = 2.0,
      opacity     = 0.9,
      popup       = ~ paste0(
        "<strong>Zone: </strong>", zone_id, "<br>",
        "<strong>Type: </strong>", zone_type
      )
    ) |>
    leaflet::setView(
      lng  = vienna_center[["lng"]],
      lat  = vienna_center[["lat"]],
      zoom = 11L
    )
} else {
  cat("No orphan zones in this snapshot — all zones contain at least one building.")
}
```

**<span class='warn'>INVESTIGATE</span> — Zones containing zero building
centroids.**

N orphan zones (this snapshot): **`r n_orphan`**.

Possible causes:

- **(a) Newly designated zone over undeveloped land** — legitimate. The zone
  designates future development requirements; no current buildings to classify.
- **(b) Inventory snapshot lag** — buildings exist physically but are not yet
  catalogued in the `GEBAEUDEOGD` inventory. The zone polygon is correct; the
  building list is incomplete.
- **(c) Zone polygon digitisation error** — the polygon was drawn in the wrong
  location. Inspect against imagery.

Investigate each orphan individually before concluding it is a data error.

### Invalid geometry

**What this is:** Buildings with self-intersecting or otherwise malformed footprint polygons as delivered by the WFS; `sf::st_make_valid()` is called before the join.
**Why it matters:** An undetected invalid geometry causes `sf::st_intersects()` to return unpredictable results and can silently exclude buildings from zone assignment.
**What to do with this:** Flag for the data provider; document each case in the audit log. Expected count in a clean run is 0.

```{r}
#| label: fig-case-invalid

sub_inv <- case_invalid(joined)

if (nrow(sub_inv) > 0L) {
  leaflet(data = sub_inv, height = "420px") |>
    leaflet::addProviderTiles(
    leaflet::providers$CartoDB.DarkMatter,
    options = leaflet::providerTileOptions(opacity = 0.95)
  ) |>
    leaflet::addCircleMarkers(
      radius      = 6,
      fillColor   = link_status_colors[["invalid_geometry"]],
      fillOpacity = 0.90,
      color       = "#8B0000",
      weight      = 1.0,
      popup       = ~ paste0(
        "<strong>Building: </strong>", building_id, "<br>",
        "<strong>Address: </strong>", address
      )
    ) |>
    leaflet::setView(
      lng  = vienna_center[["lng"]],
      lat  = vienna_center[["lat"]],
      zoom = 12L
    )
} else {
  cat(paste0(
    "No invalid-geometry buildings in this snapshot (N = 0). ",
    "st_make_valid() was applied during fetch — all footprints passed validation."
  ))
}
```

**`r if (n_invalid == 0L) "<span class='good'>PASS</span>" else "<span class='bad'>FAIL</span>"` — Self-intersecting or otherwise broken polygons.**

N invalid buildings (this snapshot): **`r n_invalid`**.

The join pipeline calls `sf::st_make_valid()` on all input geometries before
performing any spatial operations. If this count is non-zero after that step,
document the upstream data issue and report it to the data provider.

Expected in a clean run: **0**.

### CRS mismatch

**What this is:** A [coordinate reference system](https://en.wikipedia.org/wiki/Spatial_reference_system) (CRS) defines how two-dimensional map coordinates map onto the curved surface of the Earth. Two datasets can describe the same physical location using different coordinate encodings — the numbers look completely different even though the point is the same place on the ground.
**Why it matters:** A CRS mismatch between buildings and zones causes **silent ~100 m offsets** in centroid-to-zone intersection. Buildings that should be inside a zone appear outside it and vice versa — without any error message.
**What to do with this:** The pipeline detects this automatically. `join_buildings_zones()` asserts that both inputs are in [EPSG:31287](https://epsg.io/31287) before any spatial operation runs; if they differ it stops with a clear error message.

**No map for this case** — CRS errors are caught at code time, not at render time.

**CRS chain used in this POC:**

| Stage | CRS | EPSG | Purpose |
|-------|-----|------|---------|
| Native WFS data | MGI / Austria GK East | [31256](https://epsg.io/31256) | Source CRS from data provider |
| Working (analysis) | MGI / Austria Lambert | [31287](https://epsg.io/31287) | Area + distance calculations |
| Display (leaflet) | WGS84 | [4326](https://epsg.io/4326) | Leaflet rendering |

**Worked example** — the same physical building, two encodings:

```{r}
#| label: tbl-crs-worked-example
#| output: asis

# Pull a real building from joined_for_map (first row with valid geometry)
ex_row <- joined_for_map[
  !is.na(sf::st_coordinates(sf::st_centroid(joined_for_map[1,]))[1]), ][1L, ]

ex_coords_4326 <- sf::st_coordinates(sf::st_centroid(ex_row))
ex_coords_31287 <- sf::st_coordinates(
  sf::st_centroid(sf::st_transform(ex_row, crs = 31287L))
)
ex_coords_31256 <- sf::st_coordinates(
  sf::st_centroid(sf::st_transform(ex_row, crs = 31256L))
)

cat(paste0(
  "> **Worked example:** building `", ex_row$building_id[[1L]], "` ",
  "(address: ", ex_row$address[[1L]], ").\n",
  "> - [EPSG:4326](https://epsg.io/4326) (WGS84 — Leaflet): ",
  "longitude ", round(ex_coords_4326[1L], 5L), ", latitude ", round(ex_coords_4326[2L], 5L), ".\n",
  "> - [EPSG:31287](https://epsg.io/31287) (Austria Lambert — analysis): ",
  "x = ", round(ex_coords_31287[1L], 1L), " m, y = ", round(ex_coords_31287[2L], 1L), " m.\n",
  "> - [EPSG:31256](https://epsg.io/31256) (Austria GK East — native WFS): ",
  "x = ", round(ex_coords_31256[1L], 1L), " m, y = ", round(ex_coords_31256[2L], 1L), " m.\n",
  "> \n",
  "> All three rows describe the **same physical location**. ",
  "The coordinate numbers differ because each system encodes position differently — ",
  "not because the building moved."
))
```

### Snapshot lag

**What this is:** Buildings constructed after the snapshot date (`r SNAPSHOT_DATE`) cannot appear in this analysis — the WFS data was fetched on that date and reflects the inventory at that moment.
**Why it matters:** Comparing analysis results across different snapshots without accounting for the change in building inventory will produce spurious changes in all metrics.
**What to do with this:** Treat the snapshot date as the contract. If you need a more recent picture, refresh the pipeline by bumping `SNAPSHOT_DATE` in `R/snapshot.R` and re-running the fetch + join + render sequence.

**No map for this case** — temporal coverage is a data provenance issue,
not a spatial geometry issue.

**The pinned snapshot is the contract.**

Buildings constructed **after** `r SNAPSHOT_DATE` cannot appear in this
analysis regardless of their spatial location. The snapshot date is defined
once in `R/snapshot.R`:

```r
SNAPSHOT_DATE <- "`r SNAPSHOT_DATE`"
```

To refresh the analysis:

1. Bump `SNAPSHOT_DATE` in `R/snapshot.R`.
2. Re-run `fetch_acd_zones()` and `fetch_buildings()` — these write new
   Parquet files to `inst/extdata/zones_<date>.parquet` and
   `inst/extdata/buildings_<date>.parquet`.
3. Re-run `join_buildings_zones()` to produce the new joined Parquet.
4. Re-render this dashboard.

Do not compare analysis results across snapshots without accounting for the
change in building inventory.

### Distance-to-zone — the right view

**What this is:** Stage 0 found that the original cross-validation (ACD field vs spatial join) was a category error. The right way to think about zone-assignment uncertainty is the **centroid-to-boundary distance** shown below.

**Why it matters:** A building 3 m outside a zone is quite different from one 3 km away. The distance distribution gives an honest picture of how many buildings are "close calls" versus definitively outside.

```{r}
#| label: fig-distance-hist

if ("nearest_zone_dist_m" %in% names(joined)) {
  # Show distribution for outside-zone buildings only (inside buildings have dist ~0)
  outside_dist <- joined |>
    sf::st_drop_geometry() |>
    dplyr::filter(!.data$inside_any_zone | is.na(.data$inside_any_zone)) |>
    dplyr::pull(nearest_zone_dist_m)
  outside_dist <- outside_dist[!is.na(outside_dist) & outside_dist > 0]

  if (length(outside_dist) > 0L) {
    hist(
      pmin(outside_dist, 2000),  # cap at 2 km for readability
      breaks = 50,
      col    = "#9467bd",
      border = "#5c3880",
      xlab   = "Distance to nearest zone boundary (m, capped at 2,000 m)",
      ylab   = "Number of buildings",
      main   = paste0(
        "Distance to nearest zone — ", format(length(outside_dist), big.mark = ","),
        " outside-zone buildings (snapshot ", SNAPSHOT_DATE, ")"
      ),
      las = 1
    )
    abline(v = 25,  col = "#ff7f0e", lty = 2, lwd = 2)
    abline(v = 500, col = "#2ca02c", lty = 2, lwd = 2)
    legend(
      "topright",
      legend = c("25 m (Tier B/C boundary)", "500 m (Tier D/E boundary)"),
      col    = c("#ff7f0e", "#2ca02c"),
      lty    = 2, lwd = 2, bg = "#222222", text.col = "white"
    )
  } else {
    cat("No outside-zone buildings with distance data in this snapshot.")
  }
} else {
  cat("nearest_zone_dist_m column not available — histogram unavailable for this snapshot.")
}
```

The `confidence_tier` column on Page 6 thresholds this distance into 5 tiers using the constants defined in `R/join_buildings_zones.R`:

- **Tier C boundary** (orange dashed): 25 m — buildings within this distance are "just outside"
- **Tier D/E boundary** (green dashed): 500 m — buildings beyond this are far from any zone

Stage 0 finding: the original cross-validation interpreted `ACD` (a 6-digit building registry code present for ALL `r format(n_total, big.mark = ",")` buildings) as "city says this building is in a zone". This made the `acd_only` category meaningless. The correct metric is spatial distance alone.

See the full investigation: [analysis/stage_0/REPORT.md](https://github.com/JohnGavin/acd_area_climate_design/blob/main/analysis/stage_0/REPORT.md).

:::

# Stories — Worked Well / Failed Badly

## Row

::: {.panel-tabset}

### Worked

::: {.panel-tabset}

#### Story 1: The cleanly classified apartment block

```{r}
#| label: fig-story-clean

story_lat <- sf::st_coordinates(sf::st_centroid(story_clean))[, 2]
story_lng <- sf::st_coordinates(sf::st_centroid(story_clean))[, 1]

# Find the zone this building belongs to
story_zone_id <- story_clean$zone_id[[1]]
story_zone_sf <- zones[!is.na(zones$zone_id) & zones$zone_id == story_zone_id, ]

m <- leaflet(height = "400px") |>
  leaflet::addProviderTiles(
    leaflet::providers$CartoDB.DarkMatter,
    options = leaflet::providerTileOptions(opacity = 0.95)
  )

if (nrow(story_zone_sf) > 0L) {
  m <- m |>
    leaflet::addPolygons(
      data        = story_zone_sf,
      fillColor   = "#e41a1c",
      fillOpacity = 0.30,
      color       = "#e41a1c",
      weight      = 2.0,
      popup       = ~ paste0("Zone: ", zone_id)
    )
}

m |>
  leaflet::addCircleMarkers(
    data        = story_clean,
    radius      = 10,
    fillColor   = link_status_colors[["in_zone"]],
    fillOpacity = 0.95,
    color       = "#1a5276",
    weight      = 2.0,
    popup       = ~ paste0(
      "<strong>Building: </strong>", building_id, "<br>",
      "<strong>Zone: </strong>", dplyr::coalesce(zone_id, "—"), "<br>",
      "<strong>Year built: </strong>", year_built, "<br>",
      "<strong>Overlap fraction: </strong>",
      round(overlap_fraction, 3)
    )
  ) |>
  leaflet::setView(lng = story_lng, lat = story_lat, zoom = 15L)
```

Building **`r story_clean$building_id[[1]]`**, constructed
**`r story_clean$year_built[[1]]`**, falls entirely within Zone
**`r dplyr::coalesce(story_clean$zone_id[[1]], "(none)")`**
(`r dplyr::coalesce(story_clean$zone_type[[1]], "unknown")` — legal status:
`r dplyr::coalesce(story_clean$legal_status[[1]], "unknown")`).

The centroid-based join assigns this building unambiguously to one zone.
Overlap fraction: `r round(story_clean$overlap_fraction[[1]], 3)`. No manual
review required. Downstream the legal rule lookup is a simple table join on
`zone_id`.

#### Story 2: Zone-free outskirts

```{r}
#| label: fig-story-outside

outside_lat <- sf::st_coordinates(sf::st_centroid(story_outside))[, 2]
outside_lng <- sf::st_coordinates(sf::st_centroid(story_outside))[, 1]

leaflet(height = "400px") |>
  leaflet::addProviderTiles(
    leaflet::providers$CartoDB.DarkMatter,
    options = leaflet::providerTileOptions(opacity = 0.95)
  ) |>
  leaflet::addCircleMarkers(
    data        = story_outside,
    radius      = 10,
    fillColor   = link_status_colors[["outside_any_zone"]],
    fillOpacity = 0.90,
    color       = "#555555",
    weight      = 1.5,
    popup       = ~ paste0(
      "<strong>Building: </strong>", building_id, "<br>",
      "<strong>District: </strong>",
      dplyr::coalesce(district, "—"), "<br>",
      "<strong>Year built: </strong>", year_built
    )
  ) |>
  leaflet::setView(lng = outside_lng, lat = outside_lat, zoom = 14L)
```

Building **`r story_outside$building_id[[1]]`** sits in district
**`r dplyr::coalesce(story_outside$district[[1]], "unknown")`** with no ACD
zone designation. The building owner is currently free to install any heating
system — no ACD energy-planning rule binds this address as of
`r SNAPSHOT_DATE`.

This is not a data error. The building record is complete (year built:
`r story_outside$year_built[[1]]`, height: `r story_outside$height_m[[1]]` m,
address: `r story_outside$address[[1]]`). The absence of a zone assignment is
a valid, informative outcome. It means: *no ACD action required for this
property*.

:::

### Failed

::: {.panel-tabset}

#### Story 3: The straddler — half a building, half a rule

```{r}
#| label: fig-story-straddle

straddle_lat <- sf::st_coordinates(sf::st_centroid(story_straddle))[, 2]
straddle_lng <- sf::st_coordinates(sf::st_centroid(story_straddle))[, 1]

leaflet(height = "400px") |>
  leaflet::addProviderTiles(
    leaflet::providers$CartoDB.DarkMatter,
    options = leaflet::providerTileOptions(opacity = 0.95)
  ) |>
  leaflet::addPolygons(
    data        = zones,
    fillColor   = ~ zone_pal(zone_type),
    fillOpacity = 0.30,
    color       = "#333333",
    weight      = 1.5,
    popup       = ~ paste0("Zone: ", zone_id)
  ) |>
  leaflet::addCircleMarkers(
    data        = story_straddle,
    radius      = 10,
    fillColor   = link_status_colors[["straddles_boundary"]],
    fillOpacity = 0.95,
    color       = "#8B4513",
    weight      = 2.0,
    popup       = ~ paste0(
      "<strong>Building: </strong>", building_id, "<br>",
      "<strong>Primary zone: </strong>",
      dplyr::coalesce(zone_id, "—"), "<br>",
      "<strong>Overlap fraction: </strong>",
      round(overlap_fraction, 3), "<br>",
      "<strong>N overlapping zones: </strong>", n_overlapping_zones
    )
  ) |>
  leaflet::setView(lng = straddle_lng, lat = straddle_lat, zoom = 15L)
```

Building **`r story_straddle$building_id[[1]]`** sits with
**`r round(story_straddle$overlap_fraction[[1]] * 100, 0)`%** of its footprint
inside zone **`r dplyr::coalesce(story_straddle$zone_id[[1]], "(none)")`**.
The remaining
**`r round((1 - story_straddle$overlap_fraction[[1]]) * 100, 0)`%** lies
outside that zone — potentially inside another zone.

**POC limitation:** The centroid-based join assigns this building to the zone
containing its centroid. The reported `zone_id` is
**`r dplyr::coalesce(story_straddle$zone_id[[1]], "none")`** and the reported
`link_status` is `straddles_boundary` as a warning signal.

**Production fix:** Store a list-column of all overlapping zones with
individual overlap fractions. Legal teams should handle boundary-straddling
buildings case-by-case.

#### Story 4: The orphan polygon

```{r}
#| label: fig-story-orphan

if (!is.null(story_orphan) && nrow(story_orphan) > 0L) {
  orphan_lat <- sf::st_coordinates(sf::st_centroid(story_orphan))[, 2]
  orphan_lng <- sf::st_coordinates(sf::st_centroid(story_orphan))[, 1]

  leaflet(height = "400px") |>
    leaflet::addProviderTiles(
    leaflet::providers$CartoDB.DarkMatter,
    options = leaflet::providerTileOptions(opacity = 0.95)
  ) |>
    leaflet::addPolygons(
      data        = story_orphan,
      fillColor   = link_status_colors[["orphan_zone"]],
      fillOpacity = 0.40,
      color       = link_status_colors[["orphan_zone"]],
      weight      = 2.5,
      popup       = ~ paste0(
        "<strong>Zone: </strong>", zone_id, "<br>",
        "<strong>Type: </strong>", zone_type, "<br>",
        "<strong>Buildings inside: </strong>0"
      )
    ) |>
    leaflet::setView(lng = orphan_lng, lat = orphan_lat, zoom = 14L)
} else {
  cat("No orphan zones in this snapshot.")
}
```

```{r}
#| label: prose-story-orphan
#| output: asis

if (!is.null(story_orphan) && nrow(story_orphan) > 0L) {
  orphan_area <- tryCatch(
    round(
      as.numeric(sf::st_area(sf::st_transform(story_orphan, crs = 31287L))) / 10000,
      1
    ),
    error = function(e) NA_real_
  )
  cat(paste0(
    "Zone **", story_orphan$zone_id[[1]], "**",
    " (`", story_orphan$zone_type[[1]], "`)",
    " covers",
    if (!is.na(orphan_area)) paste0(" approximately **", orphan_area, " ha**") else " an unknown area",
    " but contains **zero building centroids** in the `", SNAPSHOT_DATE, "` snapshot.\n\n",
    "This could be: (a) a newly designated zone over undeveloped land —",
    " **legitimate**; (b) an inventory lag; or (c) a polygon digitisation error.",
    " Compare against Vienna satellite imagery to distinguish the cases.",
    "\n\nOrphan zones should not be conflated with the `outside_any_zone`",
    " building classification: the zone is real — it is the building list",
    " that may be incomplete relative to the zone."
  ))
} else {
  cat(paste0(
    "No orphan zones in the `", SNAPSHOT_DATE, "` snapshot — all zones contain",
    " at least one building centroid. This is a healthy result."
  ))
}
```

#### Story 5: The case of the impossible building

```{r}
#| label: prose-story-invalid
#| output: asis

if (!is.null(story_invalid) && nrow(story_invalid) > 0L) {
  cat(paste0(
    "Building **", story_invalid$building_id[[1]], "**",
    " at address **", story_invalid$address[[1]], "**",
    " has a self-intersecting footprint polygon as delivered by the WFS.\n\n",
    "After calling `sf::st_make_valid()` the geometry is repaired and the",
    " building participates in the join normally. However, we flag it as",
    " `invalid_geometry` in the output so that the data quality dashboard",
    " surfaces the upstream issue.\n\n",
    "**Why this matters:** an undetected invalid geometry would cause",
    " `sf::st_intersects()` to return unpredictable results and silently",
    " exclude the building from zone assignment. Documenting it here creates",
    " an audit trail for the data provider.\n\n",
    "N invalid in this snapshot: **", n_invalid, "**."
  ))
} else {
  cat(paste0(
    "No invalid-geometry buildings in the `", SNAPSHOT_DATE, "` snapshot.",
    " `sf::st_make_valid()` resolved all geometries before the join.",
    " This story would fire if any upstream footprint polygon is",
    " self-intersecting or otherwise malformed."
  ))
}
```

:::

:::

# How confident are we?

## Row

::: {.card title="Link Status Distribution"}

```{r}
#| label: tbl-qa-link-dist

link_summary |>
  dplyr::mutate(
    status_label = link_status_labels[as.character(link_status)]
  ) |>
  dplyr::select(status_label, n_buildings, pct) |>
  gt::gt() |>
  gt::cols_label(
    status_label = "Link status",
    n_buildings  = "N buildings",
    pct          = "%"
  ) |>
  gt::fmt_number(columns = n_buildings, decimals = 0) |>
  gt::fmt_number(columns = pct, decimals = 1) |>
  gt::data_color(
    columns = n_buildings,
    palette = c("#f7fbff", "#08306b")
  ) |>
  gt::tab_caption(
    caption = paste0(
      "Full link-status distribution. Colour-coded leaflets in earlier ",
      "tabs use the same palette (in_zone green, outside grey, etc.). ",
      "Snapshot: ", SNAPSHOT_DATE,
      ". Total buildings: ", format(n_total, big.mark = ","), "."
    )
  )
```

:::

::: {.card title="Coverage Stats"}

**Snapshot:** `r SNAPSHOT_DATE`

| Metric | Value |
|--------|-------|
| Total buildings | `r n_total` |
| Total zones | `r n_zones` |
| Buildings in any zone | `r n_total - n_outside` (`r 100 - pct_outside`%) |
| Buildings outside all zones | `r n_outside` (`r pct_outside`%) |
| Buildings straddling boundary | `r n_straddle` (`r pct_straddle`%) |
| Buildings under multiple zones | `r n_multi` (`r pct_multi`%) |
| Orphan zones (no buildings) | `r n_orphan` |
| Invalid-geometry buildings | `r n_invalid` |
| Using placeholder data | `r using_placeholder` |

:::

## Row

::: {.card title="Tier × District"}

```{r}
#| label: tbl-p6-tier-by-district

if ("confidence_tier" %in% names(joined)) {
  tier_by_dist <- tryCatch(
    acd::confidence_tier_by_district(joined),
    error = function(e) {
      # local fallback
      joined |>
        sf::st_drop_geometry() |>
        dplyr::mutate(tier = substr(as.character(.data$confidence_tier), 1L, 1L)) |>
        dplyr::count(.data$district, .data$tier, name = "n") |>
        tidyr::pivot_wider(
          names_from  = tier,
          values_from = n,
          values_fill = 0L
        )
    }
  )

  # Add focus flag and row total; sort focus districts first
  tier_letters_present <- intersect(c("A", "B", "C", "D", "E"), names(tier_by_dist))
  tier_by_dist <- tier_by_dist |>
    dplyr::mutate(
      .is_focus = .data$district %in% FOCUS_DISTRICTS_NAMES,
      .row_total = rowSums(dplyr::across(dplyr::all_of(tier_letters_present)), na.rm = TRUE)
    ) |>
    dplyr::arrange(dplyr::desc(.data$.is_focus), dplyr::desc(.data$.row_total))

  is_focus_rows <- tier_by_dist$.is_focus
  tier_by_dist <- dplyr::select(tier_by_dist, -.is_focus, -.row_total)

  tbl_tier_dist <- gt::gt(tier_by_dist) |>
    gt::cols_label(district = "District") |>
    gt::fmt_number(columns = dplyr::all_of(tier_letters_present), decimals = 0) |>
    gt::tab_style(
      style     = gt::cell_fill(color = "#fff8dc"),
      locations = gt::cells_body(rows = is_focus_rows)
    ) |>
    gt::tab_caption(
      caption = paste0(
        "Confidence tier counts by Vienna district (snapshot: ", SNAPSHOT_DATE, "). ",
        "Columns A–E are the five proximity tiers (A = inside zone, high confidence; ",
        "E = outside, far from any zone). ",
        "Focus districts Brigittenau (20) and Floridsdorf (21) are highlighted ",
        "and sorted to the top. ",
        "N = ", format(n_total, big.mark = ","), " buildings across ",
        nrow(tier_by_dist), " districts. ",
        "Source: join_buildings_zones() via ogdwien:GEBAEUDEINFOOGD + ogdwien:ENERGIEPLANERPOGD. ",
        "See STUBS.md for tier threshold constants (TIER_BOUNDARY_INSIDE_M = 5, ",
        "TIER_BOUNDARY_OUTSIDE_M = 25, TIER_NEARBY_OUTSIDE_M = 500)."
      )
    )
  tbl_tier_dist
} else {
  cat("confidence_tier column not available in this snapshot.")
}
```

:::

::: {.card title="Tier × Decade"}

```{r}
#| label: tbl-p6-tier-by-decade

if ("confidence_tier" %in% names(joined) && "year_built" %in% names(joined)) {
  tier_by_dec <- joined |>
    sf::st_drop_geometry() |>
    dplyr::filter(!is.na(.data$year_built)) |>
    dplyr::mutate(
      decade = decade(.data$year_built),
      tier   = substr(as.character(.data$confidence_tier), 1L, 1L)
    ) |>
    dplyr::count(.data$decade, .data$tier, name = "n") |>
    tidyr::pivot_wider(
      names_from  = tier,
      values_from = n,
      values_fill = 0L
    ) |>
    dplyr::arrange(dplyr::desc(.data$decade))

  tier_cols_present <- intersect(c("A", "B", "C", "D", "E"), names(tier_by_dec))

  gt::gt(tier_by_dec) |>
    gt::cols_label(decade = "Decade") |>
    gt::fmt_number(columns = dplyr::all_of(tier_cols_present), decimals = 0) |>
    gt::tab_caption(
      caption = paste0(
        "Confidence tier counts by construction decade (most recent first), ",
        "snapshot ", SNAPSHOT_DATE, ". ",
        "Decades with higher Tier A/B shares indicate districts or eras ",
        "where zone coverage is strongest. ",
        "Buildings with missing year_built are excluded from this table. ",
        "N = ", format(n_total, big.mark = ","),
        " buildings with known year_built. ",
        "Source: join_buildings_zones() via ogdwien:GEBAEUDEINFOOGD + ogdwien:ENERGIEPLANERPOGD. ",
        "Replaces the old acd_agreement by-decade table, which Stage 0 falsified ",
        "(ACD field is a building registry ID, not a zone-membership indicator)."
      )
    )
} else {
  cat("confidence_tier or year_built column not available.")
}
```

:::

::: {.card title="Sample evidence — 5 from each tier"}

```{r}
#| label: tbl-p6-sample-evidence-tier

if ("confidence_tier" %in% names(joined)) {
  set.seed(42L)

  # Centroid lon/lat for Google Maps link (joined is already WGS84)
  joined_flat <- joined |>
    sf::st_drop_geometry()

  # Add centroid coords if available in the sf object
  joined_coords <- tryCatch({
    coords <- sf::st_coordinates(sf::st_centroid(joined))
    dplyr::bind_cols(
      joined_flat,
      tibble::tibble(lon = round(coords[, 1], 6), lat = round(coords[, 2], 6))
    )
  }, error = function(e) {
    dplyr::mutate(joined_flat, lon = NA_real_, lat = NA_real_)
  })

  sample_tiers <- joined_coords |>
    dplyr::slice_sample(n = 5L, by = confidence_tier) |>
    dplyr::arrange(confidence_tier) |>
    dplyr::select(dplyr::any_of(c(
      "building_id", "district", "year_built", "address",
      "nearest_zone_dist_m", "confidence_tier", "lon", "lat"
    ))) |>
    dplyr::mutate(
      nearest_zone_dist_m = round(
        dplyr::coalesce(.data$nearest_zone_dist_m, NA_real_), 0
      ),
      google_maps = dplyr::if_else(
        !is.na(.data$lon) & !is.na(.data$lat),
        paste0("https://maps.google.com/?q=", .data$lat, ",", .data$lon),
        NA_character_
      )
    ) |>
    dplyr::select(-dplyr::any_of(c("lon", "lat")))

  if (nrow(sample_tiers) > 0L) {
    gt::gt(sample_tiers) |>
      gt::cols_label(
        building_id         = "Building ID",
        district            = "District",
        year_built          = "Year built",
        address             = "Address",
        nearest_zone_dist_m = "Dist. to zone (m)",
        confidence_tier     = "Tier",
        google_maps         = "Google Maps"
      ) |>
      gt::fmt_number(columns = dplyr::any_of("nearest_zone_dist_m"), decimals = 0) |>
      gt::fmt_url(columns = dplyr::any_of("google_maps"), label = "Map") |>
      gt::tab_caption(
        caption = paste0(
          "Random sample of 5 buildings from each confidence tier (25 rows total), ",
          "snapshot ", SNAPSHOT_DATE, ". ",
          "Columns: building_id (WFS OBJECTID), district (BEZ), year_built (BAUJAHR), ",
          "address (ADRESSE), nearest_zone_dist_m (distance in metres to nearest zone boundary ",
          "— always 0 for tiers A/B since they are inside zones), ",
          "confidence_tier (A–E), Google Maps link for visual verification. ",
          "Tiers are sorted A (highest confidence) first to E (clearly outside) last. ",
          "Source: join_buildings_zones() via ogdwien:GEBAEUDEINFOOGD + ogdwien:ENERGIEPLANERPOGD. ",
          "Use the Maps links to verify individual placements by hand."
        )
      )
  } else {
    cat("Insufficient data for tier sample table.")
  }
} else {
  cat("confidence_tier column not available in this snapshot.")
}
```

:::

## Row

::: {.card title="Drill into one zone — worked example"}

```{r}
#| label: tbl-drill-zone

# Pick an example zone: prefer focus districts (20 Brigittenau or 21 Floridsdorf),
# largest by N buildings inside; fall back to the most-populated zone overall.
if ("zone_id" %in% names(joined)) {
  zone_counts <- joined |>
    sf::st_drop_geometry() |>
    dplyr::filter(!is.na(.data$zone_id)) |>
    dplyr::count(
      .data$zone_id,
      .data$zone_district,
      .data$zone_area_ha,
      .data$zone_legal_url,
      name = "n_buildings"
    ) |>
    dplyr::mutate(.is_focus = .data$zone_district %in% c(20L, 21L)) |>
    dplyr::arrange(dplyr::desc(.data$.is_focus), dplyr::desc(.data$n_buildings))

  example_zone <- if (nrow(zone_counts) > 0L) zone_counts[1L, ] else NULL
} else {
  example_zone <- NULL
}

if (!is.null(example_zone)) {
  zone_buildings <- joined |>
    dplyr::filter(.data$zone_id == example_zone$zone_id)

  zone_n <- nrow(zone_buildings)
  zone_tier_counts <- zone_buildings |>
    sf::st_drop_geometry() |>
    dplyr::count(.data$confidence_tier, name = "n") |>
    dplyr::arrange(.data$confidence_tier)

  zone_meta <- tibble::tibble(
    Field = c("ERPLABEL", "Bezirk", "Area (ha)", "N buildings", "Legal document"),
    Value = c(
      example_zone$zone_id,
      paste0(example_zone$zone_district,
             ifelse(example_zone$zone_district %in% c(20L, 21L),
                    " (focus district)", "")),
      format(round(example_zone$zone_area_ha, 2), big.mark = ","),
      format(zone_n, big.mark = ","),
      ifelse(!is.na(example_zone$zone_legal_url) && nzchar(example_zone$zone_legal_url),
             paste0("[Verordnung](", example_zone$zone_legal_url, ")"),
             "—")
    )
  )

  gt::gt(zone_meta) |>
    gt::fmt_markdown(columns = "Value") |>
    gt::tab_caption(
      caption = paste0(
        "Worked example: zone ", example_zone$zone_id,
        " in district ", example_zone$zone_district,
        ifelse(example_zone$zone_district %in% c(20L, 21L), " (a focus district)", ""),
        ". ",
        "This is the most-populated zone among focus districts (Brigittenau, Floridsdorf); ",
        "if no focus zone has buildings, the globally most-populated zone is shown. ",
        "Use the Verordnung link to read the legal text. ",
        "Source: join_buildings_zones() via ogdwien:ENERGIERAUMPLANOGD."
      )
    )
} else {
  cat("No zone with buildings found in this snapshot.")
}
```

:::

::: {.card title="Tier breakdown for the example zone"}

```{r}
#| label: tbl-drill-zone-tiers

if (exists("zone_tier_counts") && nrow(zone_tier_counts) > 0L) {
  zone_tier_counts |>
    dplyr::mutate(
      Tier  = tier_labels[as.character(.data$confidence_tier)],
      Tier  = dplyr::coalesce(.data$Tier, as.character(.data$confidence_tier)),
      Pct   = round(100 * .data$n / sum(.data$n), 1),
      Color = tier_colors[as.character(.data$confidence_tier)]
    ) |>
    dplyr::select(Tier, .data$n, Pct, Color) |>
    dplyr::rename(`N buildings` = n) |>
    gt::gt() |>
    gt::cols_width(Color ~ gt::px(60)) |>
    gt::tab_caption(
      caption = paste0(
        "Tier distribution within zone ", example_zone$zone_id, ". ",
        "All buildings here are spatially inside the same zone polygon, so most ",
        "should be Tier A or B (deep inside / boundary inside). Any C/D/E in this ",
        "table would indicate a join anomaly. ",
        "N = ", format(zone_n, big.mark = ","), " buildings in this zone. ",
        "Source: confidence_tier column in joined data."
      )
    )
} else {
  cat("Tier breakdown unavailable.")
}
```

:::

::: {.card title="Sample buildings in the example zone"}

```{r}
#| label: tbl-drill-zone-buildings

if (exists("zone_buildings") && nrow(zone_buildings) > 0L) {
  zone_coords <- tryCatch({
    coords <- sf::st_coordinates(sf::st_centroid(zone_buildings))
    list(lon = round(coords[, 1], 6), lat = round(coords[, 2], 6))
  }, error = function(e) {
    list(lon = rep(NA_real_, nrow(zone_buildings)),
         lat = rep(NA_real_, nrow(zone_buildings)))
  })

  set.seed(42L)
  n_sample <- min(8L, nrow(zone_buildings))
  picked <- sample.int(nrow(zone_buildings), n_sample)

  sample_df <- zone_buildings[picked, ] |>
    sf::st_drop_geometry() |>
    dplyr::mutate(
      `Map link` = ifelse(
        !is.na(zone_coords$lat[picked]) & !is.na(zone_coords$lon[picked]),
        paste0("[map](https://maps.google.com/?q=",
               zone_coords$lat[picked], ",", zone_coords$lon[picked], ")"),
        "—"
      ),
      `WUKSEA-GIS` = paste0(
        "[city viewer](https://www.wien.gv.at/stadtentwicklung/energie/",
        "energieraumplanung/wuksea.html)"
      )
    ) |>
    dplyr::select(dplyr::any_of(c(
      "building_id", "address", "year_built", "n_stories",
      "nearest_zone_dist_m", "confidence_tier", "Map link", "WUKSEA-GIS"
    )))

  gt::gt(sample_df) |>
    gt::fmt_markdown(columns = c("Map link", "WUKSEA-GIS")) |>
    gt::cols_label(
      building_id         = "OBJECTID",
      address             = "Address",
      year_built          = "Built",
      n_stories           = "Storeys",
      nearest_zone_dist_m = "Dist. to boundary (m)",
      confidence_tier     = "Tier"
    ) |>
    gt::fmt_number(
      columns  = "nearest_zone_dist_m",
      decimals = 0,
      use_seps = FALSE
    ) |>
    gt::tab_caption(
      caption = paste0(
        "Random sample of ", n_sample, " buildings from zone ", example_zone$zone_id,
        " (out of ", format(zone_n, big.mark = ","), " total). ",
        "Each row links out to Google Maps for the building location and to the ",
        "city's WUKSEA-GIS viewer to verify the zone designation independently. ",
        "Use these links as a manual check on our spatial join — if the city viewer ",
        "disagrees with our classification, the building merits investigation. ",
        "Source: joined dataset, snapshot ", SNAPSHOT_DATE, "."
      )
    )
} else {
  cat("No sample buildings available.")
}
```

:::

## Row

::: {.card title="Sanity Checks"}

```{r}
#| label: tbl-sanity-checks

checks <- tibble::tibble(
  check = c(
    "All buildings have valid geometry (st_is_valid = TRUE)",
    "All zones have valid geometry (st_is_valid = TRUE)",
    "Sum of link_status percentages = 100%",
    "No NA building_id values",
    "link_status factor levels complete"
  ),
  result = c(
    if (n_invalid == 0L) "PASS" else paste0("FAIL — ", n_invalid, " invalid"),
    tryCatch({
      n_bad_z <- sum(!sf::st_is_valid(zones), na.rm = TRUE)
      if (n_bad_z == 0L) "PASS" else paste0("FAIL — ", n_bad_z, " invalid zones")
    }, error = function(e) "WARN — could not check"),
    if (abs(sum(link_summary$pct) - 100) < 0.2) "PASS" else
      paste0("FAIL — sum = ", round(sum(link_summary$pct), 1), "%"),
    if (sum(is.na(joined$building_id)) == 0L) "PASS" else
      paste0("FAIL — ", sum(is.na(joined$building_id)), " NAs"),
    if (all(
      c("in_zone", "outside_any_zone", "straddles_boundary",
        "multiple_zones", "orphan_zone", "invalid_geometry") %in%
      levels(joined$link_status)
    )) "PASS" else "FAIL — missing levels"
  )
)

checks |>
  gt::gt() |>
  gt::cols_label(check = "Check", result = "Result") |>
  gt::tab_style(
    style     = gt::cell_fill(color = "#d4edda"),
    locations = gt::cells_body(
      columns = result,
      rows    = grepl("^PASS", result)
    )
  ) |>
  gt::tab_style(
    style     = gt::cell_fill(color = "#f8d7da"),
    locations = gt::cells_body(
      columns = result,
      rows    = grepl("^FAIL", result)
    )
  ) |>
  gt::tab_style(
    style     = gt::cell_fill(color = "#fff3cd"),
    locations = gt::cells_body(
      columns = result,
      rows    = grepl("^WARN", result)
    )
  ) |>
  gt::tab_caption(caption = paste0(
    "Automated sanity checks on joined data. Snapshot: ", SNAPSHOT_DATE, "."
  ))
```

:::

::: {.card title="Recommendations"}

### What needs follow-up

1. **Straddles-boundary cases (`r n_straddle` buildings):** The POC assigns
   one zone via centroid. Production should compute area-weighted overlap and
   store a list-column of all applicable zones with fractions.

2. **Multiple-zone cases (`r n_multi` buildings):** Audit each manually to
   determine whether overlapping zone polygons are intentional (layered rules)
   or a digitisation error. If intentional, the data model must support
   multiple simultaneous zone assignments per building.

3. **Orphan zones (`r n_orphan`):** Cross-reference each orphan polygon against
   current satellite imagery and planning documents to classify as
   (a) legitimate undeveloped land, (b) inventory lag, or (c) polygon error.

4. **District assignment:** The `district` column is `NA` for all buildings in
   this POC. Production should perform a secondary spatial join against the
   Vienna district boundary layer (`BEZIRKSGRENZEOGD`) to populate this field.

### What production should add

- **Area-weighted overlap:** Replace centroid join with full footprint
  intersection to compute exact overlap fractions per zone.
- **List-column schema:** Allow multiple `(zone_id, zone_type, overlap_fraction)`
  tuples per building — required for straddlers and multi-zone buildings.
- **Address geocoding fallback:** For buildings with invalid geometry, fall back
  to address-based geocoding before marking as invalid.
- **Snapshot refresh pipeline:** Automate `SNAPSHOT_DATE` bumping and
  re-fetching via a targets pipeline target.
- **pointblank data contract:** Convert these sanity checks into a formal
  pointblank agent with alerting on refresh.

:::

## Row

::: {.card title="Remote duckplyr query (Stage 2 preview)"}

```{r}
#| label: tbl-remote-duckplyr-query

# Stage 2: demonstrate live duckplyr query against Parquet hosted on GitHub Pages.
# DuckDB issues HTTP range requests — only the link_status column is downloaded,
# not the full file. Gracefully degrades if Pages is not yet live or duckplyr
# is not available in the render environment.
demo_remote_query <- tryCatch({
  acd::load_acd_data("joined", source = "remote", collect = FALSE) |>
    dplyr::count(.data$link_status, name = "n") |>
    dplyr::collect()
}, error = function(e) {
  message("Remote duckplyr query unavailable: ", conditionMessage(e))
  NULL
})

if (!is.null(demo_remote_query)) {
  demo_remote_query |>
    gt::gt() |>
    gt::cols_label(link_status = "Link status", n = "N buildings") |>
    gt::fmt_number(columns = n, decimals = 0) |>
    gt::tab_caption(
      caption = paste0(
        "Live duckplyr query against Parquet served by GitHub Pages. ",
        "DuckDB issues HTTP range requests; only the link_status column is downloaded, ",
        "not the full file. Source URL: ",
        "https://johngavin.github.io/acd_area_climate_design/data/joined_",
        SNAPSHOT_DATE, ".parquet"
      )
    )
} else {
  cat(paste0(
    "Remote query unavailable: GitHub Pages may not yet be live, ",
    "or duckplyr is not available in this render environment. ",
    "The local snapshot (N = ", n_total, " buildings) was used throughout the rest of this dashboard."
  ))
}
```

:::

::: {.card title="Stage 2 architecture notes"}

### Parquet on GitHub Pages

The snapshot Parquet files are now committed to `docs/data/` and served by
GitHub Pages at:

`https://johngavin.github.io/acd_area_climate_design/data/`

The `load_acd_data()` function supports two modes:

| `source` | Reads from | Speed | Network |
|----------|-----------|-------|---------|
| `"local"` (default) | `inst/extdata/` (committed) | Instant | None |
| `"remote"` | GitHub Pages via DuckDB httpfs | ~1–2 s | HTTP range requests |

**Why HTTP range requests are efficient:** DuckDB reads only the byte ranges
needed for the query (predicate / projection pushdown). A `count(link_status)`
query downloads only the `link_status` column bytes — typically < 1 KB from a
30 KB Parquet file.

**Stage 3 upgrade path:** Replace `REMOTE_BASE_URL` in `R/snapshot.R` with an
HuggingFace `hf://datasets/...` or S3/R2 URL when the dataset outgrows GitHub
Pages limits (100 MB file cap, 1 GB repo cap).

:::

