---
title: "Lookup: ACD ↔ Address"
subtitle: "Find a building by ACD code or by address — Vienna's 58,255 registered buildings"
format:
  html:
    toc: false
    code-fold: true
    code-summary: "Show code"
    page-layout: full
execute:
  echo: true
  message: false
  warning: false
---

```{r}
#| label: setup
#| include: false

library(sf)
library(dplyr)
library(DT)
library(leaflet)
library(crosstalk)
library(htmltools)

# Load package in dev mode (pkgload) or installed
tryCatch(
  pkgload::load_all(quiet = TRUE),
  error = function(e) {
    tryCatch(library(acd), error = function(e2) NULL)
  }
)

if (!exists("SNAPSHOT_DATE")) SNAPSHOT_DATE <- "2026-05-01"

# Load the joined parquet (58,255 buildings)
joined_raw <- load_acd_data("joined")

# Reproject to WGS84 for Leaflet
sf_data <- sf::st_as_sf(joined_raw, wkt = "geometry_wkt", crs = CRS_WORKING) |>
  sf::st_transform(4326)

coords <- sf::st_coordinates(sf_data)

# Build the display data frame
df <- joined_raw |>
  dplyr::mutate(
    lon = coords[, 1],
    lat = coords[, 2],
    is_focus = district %in% c("20", "21"),
    Zone     = dplyr::coalesce(zone_id, "\u2014"),
    Tier     = substr(as.character(confidence_tier), 1L, 1L),
    `Distance (m)` = round(nearest_zone_dist_m, 0L)
  ) |>
  dplyr::arrange(dplyr::desc(is_focus), district, acd_native) |>
  dplyr::transmute(
    ACD            = as.character(acd_native),
    Address        = address,
    District       = district,
    `Year built`   = year_built,
    Storeys        = n_stories,
    Zone,
    `Distance (m)`,
    Tier,
    lon,
    lat
  )

# Example values for inline R (used in the quick guide below)
example_acd  <- df$ACD[df$District %in% c("20", "21")][1]
example_addr <- df$Address[df$District %in% c("20", "21")][1]
n_total      <- nrow(df)
n_focus      <- sum(df$District %in% c("20", "21"))
```

## Quick guide

* **Find one building:** paste a 6-digit ACD code such as *`r example_acd`*
  into the search box (top-right of the table) to jump straight to it.
* **Find by street:** type any street-name fragment — for example, *Praterstraße*
  or the first word of *`r example_addr`* — to see all matching buildings.
* **Drill down:** use the column-header filter row to restrict by district,
  confidence tier (A = best), or year built.
* **Focus districts:** Brigittenau (20) and Floridsdorf (21) are sorted to the
  top by default — `r n_focus` of the `r format(n_total, big.mark = ",")` buildings.
* **Map sync:** the Leaflet map below the filters updates whenever you change a
  crosstalk filter; circle colour encodes confidence tier.
* **Export:** use the CSV or Copy buttons above the table to save your selection.

---

```{r}
#| label: lookup-widget
#| echo: true

# Shared data object linking the table and map
sd <- crosstalk::SharedData$new(df)

# Colour palette for confidence tier
tier_pal <- leaflet::colorFactor(
  palette = c(
    A = "#2ecc71",
    B = "#27ae60",
    C = "#f39c12",
    D = "#e74c3c",
    E = "#8e44ad"
  ),
  domain = c("A", "B", "C", "D", "E")
)

# Table caption satisfying visualization-standards (7 items):
# 1. title/context  2. units  3. source  4. date  5. n  6. filter note  7. export note
tbl_caption <- htmltools::tags$caption(
  style = paste0(
    "caption-side:top; text-align:left; font-size:0.85em; ",
    "color:#555; padding-bottom:6px;"
  ),
  htmltools::strong("Vienna ACD building inventory — ACD \u2194 address lookup."),
  htmltools::HTML(paste0(
    " Distances in metres (rounded). Confidence tier A\u2013E: ",
    "A = building centroid inside zone, E = beyond 500\u2009m buffer. ",
    "Source: Vienna Open Data WFS, snapshot ", SNAPSHOT_DATE, ". ",
    "n = ", format(n_total, big.mark = ","), " buildings. ",
    "Crosstalk filters above also update the map. ",
    "Use CSV / Copy buttons to export the filtered selection."
  ))
)

# DT table
dt_widget <- DT::datatable(
  sd,
  caption    = tbl_caption,
  filter     = "top",
  extensions = c("Scroller", "Buttons"),
  selection  = "single",
  rownames   = FALSE,
  options    = list(
    dom        = "Bfrtip",
    buttons    = list("csv", "copy"),
    scrollY    = 500,
    scroller   = TRUE,
    pageLength = 25,
    searchHighlight = TRUE,
    # Hide lon/lat columns (index 9,10 — 0-based)
    columnDefs = list(
      list(visible = FALSE, targets = c(8L, 9L))
    )
  )
)

# Leaflet map
map_widget <- leaflet::leaflet(sd) |>
  leaflet::addProviderTiles(leaflet::providers$CartoDB.Positron) |>
  leaflet::addCircleMarkers(
    lng    = ~lon,
    lat    = ~lat,
    radius = 4,
    color  = ~tier_pal(Tier),
    stroke = FALSE,
    fillOpacity = 0.75,
    popup  = ~paste0(
      "<b>ACD:</b> ", ACD, "<br>",
      "<b>Address:</b> ", Address, "<br>",
      "<b>District:</b> ", District, "<br>",
      "<b>Year built:</b> ", `Year built`, "<br>",
      "<b>Tier:</b> ", Tier, "<br>",
      "<b>Zone:</b> ", Zone
    )
  ) |>
  leaflet::addLegend(
    position = "bottomright",
    pal      = tier_pal,
    values   = c("A", "B", "C", "D", "E"),
    title    = "Confidence tier"
  )

# Crosstalk filter panel
filter_panel <- htmltools::div(
  style = "display:flex; flex-wrap:wrap; gap:12px; margin-bottom:12px;",
  crosstalk::filter_select(
    "ct_district", "District", sd, ~District, multiple = TRUE
  ),
  crosstalk::filter_select(
    "ct_tier", "Confidence tier", sd, ~Tier, multiple = TRUE
  ),
  crosstalk::filter_slider(
    "ct_year", "Year built", sd, ~`Year built`,
    min = 1800L, max = 2025L, step = 1L
  ),
  crosstalk::filter_select(
    "ct_zone", "Zone (ACD zone ID)", sd, ~Zone, multiple = TRUE
  )
)

# Assemble: filters on top, then table | map side-by-side
htmltools::tagList(
  filter_panel,
  crosstalk::bscols(
    widths = c(7, 5),
    dt_widget,
    map_widget
  )
)
```
