---
title: "Methodology — Vienna ACD Building-Zone Linkage"
subtitle: "What we asked, what failed, what works"
format:
  html:
    toc: true
    toc-depth: 2
    code-fold: true
    code-summary: "Show code"
---

```{r}
#| label: setup
#| include: false

# Package loading — dev mode first, then installed, then graceful fail
tryCatch(
  pkgload::load_all(quiet = TRUE),
  error = function(e) {
    tryCatch(library(acd), error = function(e2) NULL)
  }
)

library(dplyr)
library(gt)
library(readr)

# Snapshot date constant (defined in R/snapshot.R; fallback if not loaded)
if (!exists("SNAPSHOT_DATE")) SNAPSHOT_DATE <- "2026-05-01"

# Load primary dataset
joined    <- load_acd_data("joined",    source = "local")
zones     <- load_acd_data("zones",     source = "local")
buildings <- load_acd_data("buildings", source = "local")

# Load OSM cross-check tables
osm_by_district <- readr::read_csv(
  system.file("extdata", "osm_coverage_by_district.csv", package = "acd"),
  show_col_types = FALSE
)
osm_total <- readr::read_csv(
  system.file("extdata", "osm_total_summary.csv", package = "acd"),
  show_col_types = FALSE
)

# Load by-district stage-0 breakdown
by_district <- readr::read_csv(
  "/private/tmp/acd_stage5/analysis/stage_0/by_district.csv",
  show_col_types = FALSE
)

# Confidence tier summary
tier_summary <- confidence_tier_summary(joined)

# Derived quantities used throughout
n_buildings  <- nrow(joined)
n_zones      <- nrow(zones)
n_districts  <- 23L  # fixed by Vienna admin structure
osm_total_n  <- osm_total$n_total[[1]]
coverage_pct <- round(100 * n_buildings / osm_total_n, 1)
```

## 1. Question

For each Vienna building in the city's open cadastre, which
[Area Climate Design (ACD)](https://www.wien.gv.at/stadtentwicklung/energie/energieraumplanung/wuksea.html)
energy-planning zone applies, and how confident are we in that linkage?

The city publishes both datasets as open
[<abbr title="Web Feature Service: OGC standard for serving geographic features over HTTP">WFS</abbr>](https://www.ogc.org/standard/wfs/)
layers, but does not provide an authoritative join key between them.
This package reconstructs the spatial linkage and quantifies confidence via proximity.

## 2. Data Sources

Three open <abbr title="Web Feature Service: OGC standard for serving geographic features over HTTP">WFS</abbr>
layers from [data.gv.at](https://www.data.gv.at/), snapshot-pinned to
`r SNAPSHOT_DATE`:

| Layer | Description | Features |
|---|---|---:|
| `ogdwien:ENERGIERAUMPLANOGD` | ACD energy-planning zones | `r n_zones` |
| `ogdwien:GEBAEUDEINFOOGD` | Building footprints with attributes | `r n_buildings` |
| `ogdwien:BEZIRKSGRENZEOGD` | District boundaries | `r n_districts` |

The zone layer contains `r n_zones` polygons covering specific development areas
across Vienna's `r n_districts` districts.
The building layer provides `r format(n_buildings, big.mark = ",")` footprints with
attributes including year of construction (`BAUJAHR`), number of floors (`GESCH_ANZ`),
and district (`BEZ`).

All three layers use
<abbr title="European Petroleum Survey Group: coordinate reference system identifier">EPSG</abbr>:31256
as delivered; the pipeline reprojects to EPSG:31287 (Austria Lambert) for spatial operations.
Schema was verified via `DescribeFeatureType` requests to the
<abbr title="Web Feature Service: OGC standard for serving geographic features over HTTP">WFS</abbr>
endpoint.

## 3. The Plan That Did Not Work

**Original approach:** compare the city's pre-linked `ACD` field on each building
against an independent spatial join.
Naive expectation: high agreement, with disagreements flagging data-quality cases.

Result on full data:

| Category | n | % |
|---|---:|---:|
| agree_in_zone | 16,246 | 27.9% |
| acd_only | 42,009 | 72.1% |
| spatial_only | 0 | 0% |
| agree_no_zone | 0 | 0% |

The 72% `acd_only` rate looked alarming: it appeared to mean 42,009 buildings where
the city says "in zone" but the spatial join says "not in zone."

The 0 `spatial_only` is also suspicious — it means every building the spatial join placed
in a zone was also flagged by the `ACD` field.

## 4. What Stage 0 Found: A Category Error

The `ACD` field in `GEBAEUDEINFOOGD` is a **building registry number**
(Adress-Code / Gebäudekennzahl), not a zone membership indicator.

Evidence:

- Every one of `r format(n_buildings, big.mark = ",")` buildings carries a unique 6-digit ACD
  value (range 1–232,762). There are zero duplicates.
- The zone layer has only `r n_zones` records. No ACD value matches any zone identifier
  across PLANNR, ERPLABEL, OBJECTID, ID, or GEBID (0 string matches; apparent numeric
  coincidences in the zero-padded space are spurious).
- Among the 16,246 buildings classified as `agree_in_zone`, none of their ACD values
  match the zone's identifier.

The cross-validation code interpreted any non-empty `ACD` field as "city says this
building is in a zone." Since all buildings have ACD populated, the code classified all
`r format(n_buildings, big.mark = ",")` buildings as "city says yes," then flagged 42,009
as `acd_only` because the spatial join placed them outside zones. This is a category error,
not a data-quality problem.

The correct reading: `r round(100 * sum(joined$inside_any_zone, na.rm = TRUE) / n_buildings, 1)`%
of Vienna buildings in this layer fall inside an ACD zone (spatial join result). The
remaining `r round(100 * (1 - sum(joined$inside_any_zone, na.rm = TRUE) / n_buildings), 1)`%
are genuinely outside all `r n_zones` zone polygons. The zone layer covers specific
development areas, not the whole city.

See `analysis/stage_0/REPORT.md` for the full investigation and hypothesis log.

## 5. What Works: Distance-Based Confidence Tiers

The defunct `acd_agreement` factor is replaced by `confidence_tier`: a 5-level factor
based on the distance from each building's centroid to the nearest zone boundary.

Computed in `R/join_buildings_zones.R` via a single
`sf::st_distance(centroids, sf::st_boundary(sf::st_union(zones)))` call.
Thresholds: 5 m (inside precision), 25 m (roughly one street width), 500 m (nearby vs
clearly outside zone coverage).

```{r}
#| label: tbl-tiers
#| tbl-cap: |
#|   **Confidence tier distribution — `r format(n_buildings, big.mark = ",")` buildings,
#|   Vienna snapshot `r SNAPSHOT_DATE`.**
#|   Tier A: centroid inside a zone, >5 m from the nearest boundary (high confidence).
#|   Tier B: inside, ≤5 m from boundary (borderline — centroid may shift with higher-resolution
#|   geometry). Tier C: outside, within 25 m of a zone edge (one street width — possible
#|   boundary misalignment). Tier D: outside, 25–500 m (zone is nearby but building is
#|   plausibly unzoned). Tier E: outside, >500 m (clearly outside any zone).
#|   Source: `R/join_buildings_zones.R`; Parquet snapshot `inst/extdata/joined_2026-05-01.parquet`.
#|   See also: §6 (per-district pattern).

tier_labels <- c(
  A_in_zone_high     = "A — in zone, high confidence",
  B_in_zone_boundary = "B — in zone, near boundary",
  C_outside_boundary = "C — outside, ≤25 m from zone",
  D_outside_nearby   = "D — outside, 25–500 m",
  E_outside_far      = "E — outside, >500 m"
)

tier_summary |>
  dplyr::mutate(
    Label      = tier_labels[as.character(confidence_tier)],
    Definition = c(
      "inside zone AND >5 m from boundary",
      "inside zone AND ≤5 m from boundary",
      "outside, ≤25 m from nearest zone",
      "outside, 25–500 m from nearest zone",
      "outside, >500 m from any zone"
    ),
    n_buildings = format(n_buildings, big.mark = ","),
    pct         = paste0(pct, "%")
  ) |>
  dplyr::select(Label, Definition, n_buildings, pct) |>
  gt::gt() |>
  gt::cols_label(
    Label       = "Tier",
    Definition  = "Distance rule",
    n_buildings = "n",
    pct         = "%"
  ) |>
  gt::cols_align(align = "right", columns = c(n_buildings, pct)) |>
  gt::tab_style(
    style = gt::cell_text(weight = "bold"),
    locations = gt::cells_column_labels()
  )
```

## 6. Per-District Pattern

Zone coverage is concentrated in the urban core and drops sharply in suburban districts.

```{r}
#| label: tbl-districts
#| tbl-cap: |
#|   **Per-district zone coverage — focus districts 20 (Brigittenau) and
#|   21 (Floridsdorf) highlighted.**
#|   n_buildings: buildings from GEBAEUDEINFOOGD in that district.
#|   pct_in_zone: percentage whose centroid falls inside an ACD zone (spatial join).
#|   Focus districts were selected for detailed study in issue #14 and show
#|   middle-ranking coverage: district 20 at
#|   `r round(100 - by_district$pct_acd_only[by_district$district == "20"], 1)`%
#|   in-zone and district 21 at
#|   `r round(100 - by_district$pct_acd_only[by_district$district == "21"], 1)`%,
#|   both well above the suburban low of `r round(100 - max(by_district$pct_acd_only), 1)`%
#|   (district 01, Innere Stadt, has the highest coverage).
#|   Source: `analysis/stage_0/by_district.csv`; spatial join via `join_buildings_zones()`.
#|   See also: §5 (confidence tiers), §7 (OSM cross-check).

district_names <- c(
  "01" = "Innere Stadt",       "02" = "Leopoldstadt",
  "03" = "Landstrasse",        "04" = "Wieden",
  "05" = "Margareten",         "06" = "Mariahilf",
  "07" = "Neubau",             "08" = "Josefstadt",
  "09" = "Alsergrund",         "10" = "Favoriten",
  "11" = "Simmering",          "12" = "Meidling",
  "13" = "Hietzing",           "14" = "Penzing",
  "15" = "Rudolfsheim-F.",     "16" = "Ottakring",
  "17" = "Hernals",            "18" = "Wahring",
  "19" = "Doebling",           "20" = "Brigittenau",
  "21" = "Floridsdorf",        "22" = "Donaustadt",
  "23" = "Liesing"
)

focus_districts <- c("20", "21")

by_district |>
  dplyr::mutate(
    pct_in_zone  = round(100 - pct_acd_only, 1),
    district_name = district_names[district],
    focus        = district %in% focus_districts
  ) |>
  dplyr::arrange(dplyr::desc(pct_in_zone)) |>
  dplyr::mutate(
    district_label = ifelse(focus,
      paste0(district, " \u2605"),
      district
    )
  ) |>
  dplyr::select(district_label, district_name, n, pct_in_zone) |>
  gt::gt() |>
  gt::cols_label(
    district_label = "District",
    district_name  = "Name",
    n              = "n buildings",
    pct_in_zone    = "% in zone"
  ) |>
  gt::tab_style(
    style = gt::cell_text(weight = "bold"),
    locations = gt::cells_body(
      rows = grepl("\u2605", district_label)
    )
  ) |>
  gt::cols_align(align = "right", columns = c(n, pct_in_zone)) |>
  gt::tab_footnote(
    footnote = "Focus districts (issue #14)",
    locations = gt::cells_column_labels(columns = district_label)
  )
```

## 7. OSM Cross-Check (Stage 2) {#sec-osm}

[OpenStreetMap (<abbr title="OpenStreetMap: the open collaborative geographic database">OSM</abbr>)](https://www.openstreetmap.org/)
provides an independent count of Vienna's building stock.

```{r}
n_osm       <- osm_total$n_total[[1]]
osm_ratio   <- round(n_osm / n_buildings, 1)
```

| Source | Total buildings |
|---|---:|
| GEBAEUDEINFOOGD (OGD) | `r format(n_buildings, big.mark = ",")` |
| <abbr title="OpenStreetMap: the open collaborative geographic database">OSM</abbr> (full Vienna bounding box) | `r format(n_osm, big.mark = ",")` |
| <abbr title="OpenStreetMap: the open collaborative geographic database">OSM</abbr> / OGD ratio | `r osm_ratio`× |

GEBAEUDEINFOOGD covers approximately
`r coverage_pct`% of Vienna's actual building stock
(`r format(n_buildings, big.mark = ",")` / `r format(n_osm, big.mark = ",")`).
This is a **curated subset**, not a comprehensive register: likely
catalogued buildings, commercial properties, and multi-residence structures
with rich administrative attributes (`BAUJAHR`, `GESCH_ANZ`, `BEZ`)
absent from the full cadastre.

This analysis covers `r coverage_pct`% of Vienna's stock. Switching to a more
complete layer is tracked in [issue #14](https://github.com/JohnGavin/acd_area_climate_design/issues/14).
The OGD layer is retained as primary because of attribute richness.

OSM data licensed under [ODbL 1.0](https://www.openstreetmap.org/copyright).
Per-district <abbr title="OpenStreetMap: the open collaborative geographic database">OSM</abbr>
counts use bounding-box queries and overcount outer districts (21, 22) due to
rectangular clip; city-wide total is authoritative.

## 8. Limitations

| Limitation | Detail |
|---|---|
| Coverage | `r coverage_pct`% of actual Vienna stock (§7) |
| Single snapshot | No time series; pinned to `r SNAPSHOT_DATE` |
| ACD field semantics | Inferred via experiments; not officially documented by the city |
| Federal cross-walk | Statistik Austria <abbr title="Gebäude- und Wohnungsregister: Austria's federal building and dwelling register">GWR</abbr> deferred (authentication friction) |
| Address validation | <abbr title="Bundesamt für Eich- und Vermessungswesen: Austria's national surveying authority">BEV</abbr> address register deferred |
| Centroid join | Not area-weighted; straddle cases simplified to primary zone |
| Multi-zone overlap | Simplified to a single primary zone per building |

## 9. How to Use the Package

```{r}
#| label: usage-parse-check
#| include: false

# Parse check per vignette-targets-export rule
usage_code <- '
library(acd)
joined <- load_acd_data("joined", source = "local")
confidence_tier_summary(joined)
confidence_tier_by_district(joined)
'
stopifnot(length(parse(text = usage_code)) > 0)
```

```r
library(acd)

# Load the full snapshot (reads from inst/extdata, no network needed)
joined <- load_acd_data("joined", source = "local")

# Confidence tier distribution
confidence_tier_summary(joined)

# District × tier pivot
confidence_tier_by_district(joined)
```

For remote access (query Parquet on GitHub Pages via DuckDB):

```r
load_acd_data("joined", source = "remote", collect = FALSE) |>
  dplyr::filter(district == "20") |>
  dplyr::count(confidence_tier) |>
  dplyr::collect()
```

## 10. Reproducibility

- Snapshot date pinned in `R/snapshot.R` (`SNAPSHOT_DATE = "r SNAPSHOT_DATE"`)
- WFS endpoints in `R/snapshot.R` (`WFS_BASE_URL`, `LAYER_ZONES`, `LAYER_BUILDINGS`, `LAYER_DISTRICTS`)
- Fetch functions (`fetch_acd_zones()`, `fetch_buildings()`, `fetch_districts()`) are cache-first
- All outputs derived from committed Parquet snapshots in `inst/extdata/`
- Nix-pinned R + system dependencies (`default.nix`)
- Code: <https://github.com/JohnGavin/acd_area_climate_design>

## 11. Links and References

- **Live dashboard:** <https://johngavin.github.io/acd_area_climate_design/vignettes/articles/dashboard.html>
- **Stage 0 investigation:** `analysis/stage_0/REPORT.md` — ACD field semantics, buffer sweep, per-district breakdown
- **Stage 2 OSM cross-check:** `analysis/stage_2/REPORT.md` — full methodology for OSM comparison
- **Issue #14:** GEBAEUDEINFOOGD vs full building stock (coverage investigation)
- **Issue #15:** Methodology vignette (this document)
- **City's official tool:** [WUKSEA-GIS](https://www.wien.gv.at/stadtentwicklung/energie/energieraumplanung/wuksea.html) — interactive ACD zone viewer
- **OSM data:** © [OpenStreetMap contributors](https://www.openstreetmap.org/copyright), ODbL 1.0

---

## Recent Changes

<!-- Placeholder: updated by each release -->
- 2026-05-01: Stage 5 — methodology vignette added (closes #15)
